---
title: "Agreement of data sources by eGFR"
author: "Shalini Santhakumaran"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    theme: null
    highlight: null
    mathjax: null
    toc: false
    fig_caption: false
    df_print: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Load libraries
library(tidyverse)
library(knitr)
library(kableExtra)

# Create directory if it doesnt exist
fs::dir_create(here::here("output", "notebooks"))

#read in dataset
#pick up 1st Jan cohort as latest renal status is on or before index date
 df<-read.csv(here::here("output", "ukrr", "input_ukrr_2020-01-01.csv.gz")) 

#subset variable to reduce memory demands
df<-df[,c("ukrr_2019", "ukrr_2019_mod","latest_rrt_status", "latest_rrt_status_secondary", "ckd_primis_stage", 
        "ckd_egfr_category", "egfr_numeric_value_history", "egfr_numeric_value_history_operator",
          "ip_ckd1_diagnosis_date", "ip_ckd2_diagnosis_date", "ip_ckd3_diagnosis_date","ip_ckd4_diagnosis_date", "ip_ckd5_diagnosis_date")] 
 

# add variable for eGFR based on last value
df<- df %>%
  mutate(ckd_egfr_category_lastvalue = case_when (
    egfr_numeric_value_history >= 90 &
      ((! egfr_numeric_value_history_operator == "<" &
          ! egfr_numeric_value_history_operator == "<=") |
         is.na(egfr_numeric_value_history_operator)) ~ "G1" ,
    egfr_numeric_value_history >= 60 &
      egfr_numeric_value_history < 90 &
      ((! egfr_numeric_value_history_operator == "<" &
          ! egfr_numeric_value_history_operator == "<=" &
          ! egfr_numeric_value_history_operator == ">" &
          ! egfr_numeric_value_history_operator == ">=" ) |
         is.na(egfr_numeric_value_history_operator)) ~ "G2" ,
    egfr_numeric_value_history >= 45 &
      egfr_numeric_value_history < 60 &
      ((! egfr_numeric_value_history_operator == "<" &
          ! egfr_numeric_value_history_operator == "<=" &
          ! egfr_numeric_value_history_operator == ">" &
          ! egfr_numeric_value_history_operator == ">=" )|
         is.na(egfr_numeric_value_history_operator)) ~ "G3a" ,
    egfr_numeric_value_history >= 30 &
      egfr_numeric_value_history < 45 &
      ((! egfr_numeric_value_history_operator == "<" &
          ! egfr_numeric_value_history_operator == "<=" &
          ! egfr_numeric_value_history_operator == ">" &
          ! egfr_numeric_value_history_operator == ">=" )|
         is.na(egfr_numeric_value_history_operator)) ~ "G3b" ,
    egfr_numeric_value_history >= 15 &
      egfr_numeric_value_history < 30 &
      ((! egfr_numeric_value_history_operator == "<" &
          ! egfr_numeric_value_history_operator == "<=" &
          ! egfr_numeric_value_history_operator == ">" &
          ! egfr_numeric_value_history_operator == ">="  )|
         is.na(egfr_numeric_value_history_operator)) ~ "G4" ,
    egfr_numeric_value_history > 0 &
      egfr_numeric_value_history < 15 &
      ((! egfr_numeric_value_history_operator == ">" &
          ! egfr_numeric_value_history_operator == ">=" )|
         is.na(egfr_numeric_value_history_operator)) ~ "G5" 
  ))

  
# add variable for eGFR based on last value, but assuming the highest CKD stage from the numeric values when there is a > or >= operator (not many with < or  <= operator)
# so if a person has >60 they will be stage G2 but becuase of the category boundaries
df<- df %>%
  mutate(ckd_egfr_category_operator = case_when (
    egfr_numeric_value_history >= 90 &
      ((! egfr_numeric_value_history_operator == "<" &
          ! egfr_numeric_value_history_operator == "<=") |
         is.na(egfr_numeric_value_history_operator)) ~ "G1" ,
    egfr_numeric_value_history >= 60 &
      egfr_numeric_value_history < 90 &
      ((! egfr_numeric_value_history_operator == "<" &
          ! egfr_numeric_value_history_operator == "<=" ) |
         is.na(egfr_numeric_value_history_operator)) ~ "G2" ,
    egfr_numeric_value_history >= 45 &
      egfr_numeric_value_history < 60 &
      ((! egfr_numeric_value_history_operator == "<" &
          ! egfr_numeric_value_history_operator == "<="  )|
         is.na(egfr_numeric_value_history_operator)) ~ "G3a" ,
    egfr_numeric_value_history >= 30 &
      egfr_numeric_value_history < 45 &
      ((! egfr_numeric_value_history_operator == "<" &
          ! egfr_numeric_value_history_operator == "<="  )|
         is.na(egfr_numeric_value_history_operator)) ~ "G3b" ,
    egfr_numeric_value_history >= 15 &
      egfr_numeric_value_history < 30 &
      ((! egfr_numeric_value_history_operator == "<" &
          ! egfr_numeric_value_history_operator == "<="  )|
         is.na(egfr_numeric_value_history_operator)) ~ "G4" ,
    egfr_numeric_value_history > 0 &
      egfr_numeric_value_history < 15  ~ "G5" 
  ))

#add variable for dialysis/tx for UKRR
df<- df %>%
  mutate(ukrr_dtx = ifelse(ukrr_2019_mod %in% c("ICHD", "HD", "HHD", "PD") , "Dialysis", ifelse(ukrr_2019_mod == "Tx", "Transplant", "Not on RRT")))
# order factor
df$ukrr_dtx<-factor(df$ukrr_dtx, levels=c("Not on RRT", "Dialysis", "Transplant"))

# add variable for source
df <- df %>%
  mutate(source_prev_rrt = case_when(ukrr_2019==1 & 
                                      latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "UPS",
                                      ukrr_2019!=1 & 
                                      latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "PS",
                                     ukrr_2019==1 & 
                                      !(latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown")) &
                                      latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "US",
                                     ukrr_2019==1 & 
                                      latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      !(latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown")) ~ "UP",
                                     ukrr_2019==1 & 
                                      !(latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown")) &
                                      !(latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown")) ~ "U",
                                     ukrr_2019!=1 & 
                                      latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      !(latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown")) ~ "P",
                                     ukrr_2019!=1 & 
                                      !(latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown")) &
                                      latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "S")) %>%
  mutate(source_prev_dialysis = case_when(ukrr_dtx=="Dialysis" & 
                                      latest_rrt_status  =="Dialysis" &
                                      latest_rrt_status_secondary =="Dialysis" ~ "UPS",
                                      ukrr_dtx!="Dialysis" & 
                                      latest_rrt_status  =="Dialysis" &
                                      latest_rrt_status_secondary =="Dialysis" ~ "PS",
                                     ukrr_dtx=="Dialysis" & 
                                      !(latest_rrt_status  =="Dialysis") &
                                      latest_rrt_status_secondary =="Dialysis" ~ "US",
                                     ukrr_dtx=="Dialysis" & 
                                      latest_rrt_status  =="Dialysis" &
                                      !(latest_rrt_status_secondary =="Dialysis") ~ "UP",
                                     ukrr_dtx=="Dialysis" & 
                                      !(latest_rrt_status  =="Dialysis") &
                                      !(latest_rrt_status_secondary =="Dialysis") ~ "U",
                                     ukrr_dtx!="Dialysis" & 
                                      latest_rrt_status  =="Dialysis" &
                                      !(latest_rrt_status_secondary =="Dialysis") ~ "P",
                                     ukrr_dtx!="Dialysis" & 
                                      !(latest_rrt_status  =="Dialysis") &
                                      latest_rrt_status_secondary =="Dialysis" ~ "S")) %>%  
    mutate(source_prev_transplant = case_when(ukrr_dtx=="Transplant" & 
                                      latest_rrt_status  =="Transplant" &
                                      latest_rrt_status_secondary =="Transplant" ~ "UPS",
                                      ukrr_dtx!="Transplant" & 
                                      latest_rrt_status  =="Transplant" &
                                      latest_rrt_status_secondary =="Transplant" ~ "PS",
                                     ukrr_dtx=="Transplant" & 
                                      !(latest_rrt_status  =="Transplant") &
                                      latest_rrt_status_secondary =="Transplant" ~ "US",
                                     ukrr_dtx=="Transplant" & 
                                      latest_rrt_status  =="Transplant" &
                                      !(latest_rrt_status_secondary =="Transplant") ~ "UP",
                                     ukrr_dtx == "Transplant" & 
                                      !(latest_rrt_status  =="Transplant") &
                                      !(latest_rrt_status_secondary =="Transplant") ~ "U",
                                     ukrr_dtx!="Transplant" & 
                                      latest_rrt_status  =="Transplant" &
                                      !(latest_rrt_status_secondary =="Transplant") ~ "P",
                                     ukrr_dtx!="Transplant" & 
                                      !(latest_rrt_status  =="Transplant") &
                                      latest_rrt_status_secondary =="Transplant" ~ "S")) 



# order factor
source_levels<-c("UPS", "UP", "US", "PS", "U", "P", "S")
df$source_prev_rrt<-factor(df$source_prev_rrt, levels=source_levels)


#calculate latest CKD stage in secondary care (the variable latest_ckd_status_secondary includes RRT which we don't want here)
#we have the dates of each ICD10 code for the CKD stages, so want to pick the latest one
#first get the dates of the ckd stages in secondary care

ckd_secondary_dates1<-df[,c("ip_ckd1_diagnosis_date", "ip_ckd2_diagnosis_date", "ip_ckd3_diagnosis_date",
                           "ip_ckd4_diagnosis_date", "ip_ckd5_diagnosis_date")]  %>%
  mutate_all( function(x) as.Date(x))  #convert to date

#vector of where there are no CKD codes at all
no_ckd<-rowSums(is.na(ckd_secondary_dates1))==5

#making changes needed to use max.col function
ckd_secondary_dates2<-ckd_secondary_dates1 %>%
         replace(is.na(.), as.Date("1900-01-01")) %>%  # replace missing with earliest date for max.col
          mutate_all(function(x) as.numeric(x))  #convert to numeric for max.col

#get which column has the latest date - the fist col is stage 1, second is stage 2 etc
ckd_stage_secondary<-max.col(ckd_secondary_dates2, ties.method="last")
#repalce with missing when all there is no ckd code (as earlier we replace the missings with an early date "1900-01-01")
ckd_stage_secondary[no_ckd]<-NA

#setting as a column
df$ckd_stage_sec<-ckd_stage_secondary


#apply disclosure control - redact if less than 7, then round to nearest 5
disclosure <-function(x, whichmargins=0){
  redact <- 0<x & x<=7  #flag redacted values for later
  x[redact] <- 0  # set redacted to 0
  x<-round(x/5)*5 # round to nearest 5
  if(whichmargins!=0){
    x<-addmargins(x,whichmargins) #add totals
  }
  if(whichmargins==1){
    redact<-rbind(redact, rep(FALSE,ncol(redact))) #add extra row to redaction, otherwise won't blank correctly
  }
  if(whichmargins==2){
    redact<-cbind(redact, rep(FALSE,nrow(redact))) #add extra col to redaction, otherwise won't blank correctly
  }
  x[redact]<-"<=7"  # flag redacted values
  return(x)
}
```

 
## Numbers needed to calculate sensitivity and specificity by modality

Agreement is indicated by the data source, defined as UPS = in the **U**KRR, **P**rimary care and **S**econdary care data; UP = in **U**KRR and **P**rimary care only; U = in **U**KRR only etc.

### Dialysis, eGFR<15
```{r, echo=FALSE, results='asis', error=TRUE}
disclosure(table(df$source_prev_dialysis[df$ckd_egfr_category_lastvalue=="G5"], useNA="ifany"),0) %>%
  kbl(col.names=c("Source", "N")) %>%
  kable_styling()
```
### Transplant, eGFR<15
```{r, echo=FALSE, results='asis', error=TRUE}
disclosure(table(df$source_prev_transplant[df$ckd_egfr_category_lastvalue=="G5"], useNA="ifany"),0) %>%
  kbl(col.names=c("Source", "N")) %>%
  kable_styling()
```


### Dialysis, eGFR>=15 or missing
```{r, echo=FALSE, results='asis', error=TRUE}
disclosure(table(df$source_prev_dialysis[(df$ckd_egfr_category_lastvalue!="G5" | is.na(df$ckd_egfr_category_lastvalue)) ], useNA="ifany"),0) %>%
  kbl(col.names=c("Source", "N")) %>%
  kable_styling()
```
### Transplant, eGFR>=15 or missing
```{r, echo=FALSE, results='asis', error=TRUE}
disclosure(table(df$source_prev_transplant[(df$ckd_egfr_category_lastvalue!="G5" | is.na(df$ckd_egfr_category_lastvalue)) ], useNA="ifany"),0) %>%
  kbl(col.names=c("Source", "N")) %>%
  kable_styling()
```


## CKD staging by eGFR (last only) in PC for people on KRT in PC and also UKRR, by PC and UKRR modality  
```{r, echo=FALSE, results='asis', error=TRUE}
k<-disclosure(table(df[df$source_prev_rrt %in% c("UP", "UPS"),]$latest_rrt_status,
       df[df$source_prev_rrt %in% c("UP", "UPS"),]$ckd_egfr_category_lastvalue, 
      df[df$source_prev_rrt %in% c("UP", "UPS"),]$ukrr_dtx, useNA = "always")  )
k2<-k%>%
  kbl() %>%
  add_header_above(c("PC status", "CKD stage", "UKRR status", "Freq")) %>%
  kable_styling()
print(k2)
```