---
title: "Checking secondary care data"
author: "Shalini Santhakumaran"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

The number of people on and starting RRT according to secondary care data seems too high. When selecting admissions/appointments based on a codelist it is not possible to return the code used, so a separate study definition was used, creating variables for each code for just 3 months of data as a check.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Load libraries
library(tidyverse)
library(knitr)
library(kableExtra)

### We also check the incident start dates of the people in the main study definition but not doing this for now to get some numbers before study def finishes.

# Create directory if it doesnt exist
fs::dir_create(here::here("output", "notebooks"))

#read in datasets
#pick up 1st Jan cohort as latest renal status is on or before index date
df_check<-read.csv(here::here("output",  "input_check_SC.csv.gz"))
#using the ethnicity study def as main study def is running -can still get incidence variables so doesn't matter
df_main<-read.csv(here::here("output", "input_ukrr_ethnicity.csv.gz"))
 
#list of secondary care cod variables 
SC_vars<-names(df_check)[grepl('proc_|diag_', names(df_check))]
#list of variables with critical care info
CC_vars<-names(df_check)[grepl('_cc_', names(df_check))]
#create flag for cc vars
CC_flags<-df_check[CC_vars]
CC_flags[CC_flags>0]<-1

demogs<-c("age_band", "sex", "ethnicity", "imd", "diabetes", "hypertension",  "at_risk", "region")
ukrr<-c("ukrr_2019", "ukrr_2020", "ukrr_2021", "ukrr_ckd2020", "ukrr_inc2020")
 
 #getting year of start for incident start dates
inc_year_pc<-as.numeric(substr(df_main$incident_rrt_date,1,4))
inc_year_sc<-as.numeric(substr(df_main$incident_rrt_date_secondary,1,4))
#ukrr - start dates are stored in different variables, start with inc 2020 then look at prevalent cohorts, replacing missing values only
#will be consistent as we only have one start date per patient in the data that was sent
inc_year_ukrr<-as.numeric(substr(df_main$ukrr_inc2020_date,1,4))
inc_year_ukrr[is.na(inc_year_ukrr)]<-as.numeric(substr(df_main$ukrr_2019_startdate,1,4))
inc_year_ukrr[is.na(inc_year_ukrr)]<-as.numeric(substr(df_main$ukrr_2020_startdate,1,4))
inc_year_ukrr[is.na(inc_year_ukrr)]<-as.numeric(substr(df_main$ukrr_2021_startdate,1,4))
 
 #looking at whether people have any of the indivudal codes, or any codes on the codelists. 
 #This is to check that the individual code variables are consistent with the study def variables that look at the whole codelist together.
 any_individual<-apply(df_check[SC_vars], 1, max)
any_list<-apply(df_check[c("op_RRT", "ip_RRT_diagnosis", "ip_RRT_procedure")], 1, max)


#function to round up to nearest accuracy
roundup = function(x, accuracy, f=ceiling){f(x/ accuracy) * accuracy}
```

Tablulate all the binary flags for procedure and diagnosis secondary care codes

```{r, echo=FALSE, results='asis', error=TRUE}

t<- t(as.matrix(apply(df_check[SC_vars], 2, function(x) table(factor(x, levels=0:1)))))
  k<- t  %>%
          kbl( booktabs=TRUE) %>%
          kable_styling()
  print(k)
```


Compare patients who have any of the individual flags with the overall RRT flag- shouldn be the same

```{r, echo=FALSE, results='asis', error=TRUE}

t<- table(any_individual,any_list)

  k<- t %>%
          kbl( booktabs=TRUE) %>%
          add_header_above(c("Overall RRT"=ncol(t)+1)) %>%
          kable_styling()
  print(k)
```

Look at which codes appear based on whether they are in the UKRR or not
```{r, echo=FALSE, results='asis', error=TRUE}

list.of.dts<-lapply(df_check[SC_vars], function(x) table(x, df_check$ukrr_2019, useNA="ifany"))

for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=SC_vars[i], booktabs=TRUE) %>%
          add_header_above(c("Overall RRT"=ncol(t)+1)) %>%
          kable_styling()
  print(k)
}
```

Look at which codes have any critical care days - need to compare this with total for that procedure/diagnosis
```{r, echo=FALSE, results='asis', error=TRUE}

t<- t(as.matrix(apply(CC_flags, 2, function(x) table(factor(x, levels=0:1)))))
  k<- t  %>%
          kbl( booktabs=TRUE) %>%
          kable_styling()
  print(k)
```

Incident start dates in secondary care from main cohort

```{r, echo=FALSE, results='asis', error=TRUE}
t<-table(inc_year_sc, inc_year_pc, useNA="ifany")
  k<- t %>%
          kbl( booktabs=TRUE) %>%
          add_header_above(c("Inc year SC", "Inc year PC"=ncol(t))) %>%
          kable_styling()
  print(k)
  
t<-table(inc_year_sc, inc_year_ukrr, useNA="ifany")
  k<- t %>%
          kbl( booktabs=TRUE) %>%
          add_header_above(c("Inc year SC", "Inc year UKRR"=ncol(t))) %>%
          kable_styling()
  print(k)
  
t<-table(inc_year_pc, inc_year_ukrr, useNA="ifany")
  k<- t %>%
          kbl( booktabs=TRUE) %>%
          add_header_above(c("Inc year PC", "Inc year UKRR"=ncol(t))) %>%
          kable_styling()
  print(k)

```
NOt sure why we aren't pickng up as many incident patients in PC - we should have start dates for all prevalent patients
at least because they have at least one RRT entry.Comparing incident flags with and without the ignore_missing_values

```{r, echo=FALSE, results='asis', error=TRUE}

t<-table(df_check$earliest_RRT, df_check$earliest_RRT_b, useNA="ifany")
  k<- t %>%
          kbl( caption="RRT flag", booktabs=TRUE) %>%
          kable_styling()
  print(k)

t<-table(as.numeric(substr(df_check$earliest_RRT_date,1,4)), as.numeric(substr(df_check$earliest_RRT_b_date,1,4)), useNA="ifany")
  k<- t %>%
          kbl( caption="RRT dates", booktabs=TRUE) %>%
          kable_styling()
  print(k)

t<-table(df_check$earliest_dialysis, df_check$earliest_dialysis_b, useNA="ifany")
  k<- t %>%
          kbl( caption="dialysis flag", booktabs=TRUE) %>%
          kable_styling()
  print(k)
  
t<-table(as.numeric(substr(df_check$earliest_dialysis_date,1,4)), as.numeric(substr(df_check$earliest_dialysis_b_date,1,4)), useNA="ifany")
  k<- t %>%
          kbl( caption="dialysis dates", booktabs=TRUE) %>%
          kable_styling()
  print(k)

  t<-table(df_check$earliest_kidney_tx, df_check$earliest_kidney_tx_b, useNA="ifany")
  k<- t %>%
          kbl( caption="kidney_tx flag", booktabs=TRUE) %>%
          kable_styling()
  print(k)

t<-table(as.numeric(substr(df_check$earliest_kidney_tx_date,1,4)), as.numeric(substr(df_check$earliest_kidney_tx_b_date,1,4)), useNA="ifany")
  k<- t %>%
          kbl( caption="kidney_tx dates", booktabs=TRUE) %>%
          kable_styling()
  print(k)

```

Adding comparison with prevalent - should have an earliest date for everyone with a latest date.

```{r, echo=FALSE, results='asis', error=TRUE}

t<-table(df_check$earliest_RRT, df_check$earliest_RRT_b, useNA="ifany")
  k<- t %>%
          kbl( caption="RRT flag", booktabs=TRUE) %>%
          kable_styling()
  print(k)

t<-table(as.numeric(substr(df_check$earliest_RRT_date,1,4)), as.numeric(substr(df_check$earliest_RRT_b_date,1,4)), useNA="ifany")
  k<- t %>%
          kbl( caption="RRT dates", booktabs=TRUE) %>%
          kable_styling()
  print(k)

t<-table(df_check$earliest_dialysis, df_check$earliest_dialysis_b, useNA="ifany")
  k<- t %>%
          kbl( caption="dialysis flag", booktabs=TRUE) %>%
          kable_styling()
  print(k)
  
t<-table(as.numeric(substr(df_check$earliest_dialysis_date,1,4)), as.numeric(substr(df_check$earliest_dialysis_b_date,1,4)), useNA="ifany")
  k<- t %>%
          kbl( caption="dialysis dates", booktabs=TRUE) %>%
          kable_styling()
  print(k)

  t<-table(df_check$earliest_kidney_tx, df_check$earliest_kidney_tx_b, useNA="ifany")
  k<- t %>%
          kbl( caption="kidney_tx flag", booktabs=TRUE) %>%
          kable_styling()
  print(k)

t<-table(as.numeric(substr(df_check$earliest_kidney_tx_date,1,4)), as.numeric(substr(df_check$earliest_kidney_tx_b_date,1,4)), useNA="ifany")
  k<- t %>%
          kbl( caption="kidney_tx dates", booktabs=TRUE) %>%
          kable_styling()
  print(k)

```
