---
title: "Checking secondary care data"
author: "Shalini Santhakumaran"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

The number of people on and starting RRT according to secondary care data seems too high. When selecting admissions/appointments based on a codelist it is not possible to return the code used, so a separate study definition was used, creating variables for each code for just 3 months of data as a check. We also check the incident start dates of the people in the main study definition.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Load libraries
library(tidyverse)
library(knitr)
library(kableExtra)

# Create directory if it doesnt exist
fs::dir_create(here::here("output", "notebooks"))

#read in datasets
#pick up 1st Jan cohort as latest renal status is on or before index date
df_check<-read.csv(here::here("output",  "input_check_SC.csv.gz"))
 df_main<-read.csv(here::here("output", "joined", "input_2020-01-01.csv.gz"))
 
SC_vars<-names(df_check)[grepl('proc_|diag_', names(df_check))]
demogs<-c("age_band", "sex", "ethnicity", "imd", "diabetes", "hypertension",  "at_risk", "region")
ukrr<-c("ukrr_2019", "ukrr_2020", "ukrr_2021", "ukrr_ckd2020", "ukrr_inc2020")
 
 #getting year of start for incident start dates
 inc_year_pc<-as.numeric(substr(df_main$incident_rrt_date,1,4))
 inc_year_sc<-as.numeric(substr(df_main$incident_rrt_date_secondary,1,4))
 
 #looking at whether people have any of the indivudal codes, or any codes on the codelists. 
 #This is to check that the individual code variables are consistent with the study def variables that look at the whole codelist together.
 any_individual<-apply(df_check[SC_vars], 1, max)
any_list<-apply(df_check[c("op_RRT", "ip_RRT_diagnosis", "ip_RRT_procedure")], 1, max)

#function to round up to nearest accuracy
roundup = function(x, accuracy, f=ceiling){f(x/ accuracy) * accuracy}
```

Tablulate all the binary flags for procedure and diagnosis secondary care codes

```{r, echo=FALSE, results='asis', error=TRUE}

t<- t(as.matrix(apply(df_check[SC_vars], 2, table)))
  k<- t  %>%
          kbl( booktabs=TRUE) %>%
          kable_styling()
  print(k)
```

Compare patients who have any of the individual flags with the overall RRT flag- shouldn be the same

```{r, echo=FALSE, results='asis', error=TRUE}

t<- table(any_individual,any_list)

  k<- t %>%
          kbl( booktabs=TRUE) %>%
          add_header_above(c("Overall RRT"=ncol(t))) %>%
          kable_styling()
  print(k)
```

Look at which codes appear based on whether they are in the UKRR or not


```{r, echo=FALSE, results='asis', error=TRUE}

list.of.dts<-lapply(df_check[SC_vars], function(x) table(x, df_check$ukrr_2019, useNA="ifany"))

for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=SC_vars[i], booktabs=TRUE) %>%
          add_header_above(c("Overall RRT"=ncol(t)+1)) %>%
          kable_styling()
  print(k)
}
```

Incident start dates in secondary care from main cohort

```{r, echo=FALSE, results='asis', error=TRUE}
t<-table(inc_year_sc, inc_year_pc)
  k<- t %>%
          kbl( booktabs=TRUE) %>%
          add_header_above(c("Inc year SC", "Inc year PC"=ncol(t))) %>%
          kable_styling()
  print(k)
  
list.of.dts<-lapply(df_main[ukrr], function(x) table(inc_year_sc,x, useNA="ifany"))

for(i in 1:length(list.of.dts)) {
  t<-list.of.dts[[i]]
  k<- t %>%
          kbl( caption = ukrr[i], booktabs=TRUE) %>%
          add_header_above(c("Inc year SC", "ukrr" =ncol(t)  )) %>%
          kable_styling()
  print(k)
}
```
