---
title: "Checking CKD stage with eGFR"
author: "Shalini Santhakumaran"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

This document describes the latest CKD stage of people who had a KRT code in primary or secondary care but are not in the UKRR data, first using CKD codes and then (for primary care only) eGFR data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Load libraries
library(tidyverse)
library(knitr)
library(kableExtra)

# Create directory if it doesnt exist
fs::dir_create(here::here("output", "notebooks"))

#read in dataset
#pick up 1st Jan cohort as latest renal status is on or before index date
 df<-read.csv(here::here("output", "ukrr", "input_ukrr_2020-01-01.csv.gz")) 

#subset variable to reduce memory demands
df<-df[,c("ukrr_2019", "latest_rrt_status", "latest_rrt_status_secondary", "ckd_primis_stage", 
        "ckd_egfr_category", "egfr_numeric_value_history", "egfr_numeric_value_history_operator",
          "ip_ckd1_diagnosis_date", "ip_ckd2_diagnosis_date", "ip_ckd3_diagnosis_date","ip_ckd4_diagnosis_date", "ip_ckd5_diagnosis_date")] 
 

# add variable for eGFR based on last value
df<- df %>%
  mutate(ckd_egfr_category_lastvalue = case_when (
    egfr_numeric_value_history >= 90 &
      ((! egfr_numeric_value_history_operator == "<" &
          ! egfr_numeric_value_history_operator == "<=") |
         is.na(egfr_numeric_value_history_operator)) ~ "G1" ,
    egfr_numeric_value_history >= 60 &
      egfr_numeric_value_history < 90 &
      ((! egfr_numeric_value_history_operator == "<" &
          ! egfr_numeric_value_history_operator == "<=" &
          ! egfr_numeric_value_history_operator == ">" &
          ! egfr_numeric_value_history_operator == ">=" ) |
         is.na(egfr_numeric_value_history_operator)) ~ "G2" ,
    egfr_numeric_value_history >= 45 &
      egfr_numeric_value_history < 60 &
      ((! egfr_numeric_value_history_operator == "<" &
          ! egfr_numeric_value_history_operator == "<=" &
          ! egfr_numeric_value_history_operator == ">" &
          ! egfr_numeric_value_history_operator == ">=" )|
         is.na(egfr_numeric_value_history_operator)) ~ "G3a" ,
    egfr_numeric_value_history >= 30 &
      egfr_numeric_value_history < 45 &
      ((! egfr_numeric_value_history_operator == "<" &
          ! egfr_numeric_value_history_operator == "<=" &
          ! egfr_numeric_value_history_operator == ">" &
          ! egfr_numeric_value_history_operator == ">=" )|
         is.na(egfr_numeric_value_history_operator)) ~ "G3b" ,
    egfr_numeric_value_history >= 15 &
      egfr_numeric_value_history < 30 &
      ((! egfr_numeric_value_history_operator == "<" &
          ! egfr_numeric_value_history_operator == "<=" &
          ! egfr_numeric_value_history_operator == ">" &
          ! egfr_numeric_value_history_operator == ">="  )|
         is.na(egfr_numeric_value_history_operator)) ~ "G4" ,
    egfr_numeric_value_history > 0 &
      egfr_numeric_value_history < 15 &
      ((! egfr_numeric_value_history_operator == ">" &
          ! egfr_numeric_value_history_operator == ">=" )|
         is.na(egfr_numeric_value_history_operator)) ~ "G5" 
  ))

  
# add variable for eGFR based on last value, but assuming the highest CKD stage from the numeric values when there is a > or >= operator (not many with < or  <= operator)
# so if a person has >60 they will be stage G2 but becuase of the category boundaries
df<- df %>%
  mutate(ckd_egfr_category_operator = case_when (
    egfr_numeric_value_history >= 90 &
      ((! egfr_numeric_value_history_operator == "<" &
          ! egfr_numeric_value_history_operator == "<=") |
         is.na(egfr_numeric_value_history_operator)) ~ "G1" ,
    egfr_numeric_value_history >= 60 &
      egfr_numeric_value_history < 90 &
      ((! egfr_numeric_value_history_operator == "<" &
          ! egfr_numeric_value_history_operator == "<=" ) |
         is.na(egfr_numeric_value_history_operator)) ~ "G2" ,
    egfr_numeric_value_history >= 45 &
      egfr_numeric_value_history < 60 &
      ((! egfr_numeric_value_history_operator == "<" &
          ! egfr_numeric_value_history_operator == "<="  )|
         is.na(egfr_numeric_value_history_operator)) ~ "G3a" ,
    egfr_numeric_value_history >= 30 &
      egfr_numeric_value_history < 45 &
      ((! egfr_numeric_value_history_operator == "<" &
          ! egfr_numeric_value_history_operator == "<="  )|
         is.na(egfr_numeric_value_history_operator)) ~ "G3b" ,
    egfr_numeric_value_history >= 15 &
      egfr_numeric_value_history < 30 &
      ((! egfr_numeric_value_history_operator == "<" &
          ! egfr_numeric_value_history_operator == "<="  )|
         is.na(egfr_numeric_value_history_operator)) ~ "G4" ,
    egfr_numeric_value_history > 0 &
      egfr_numeric_value_history < 15  ~ "G5" 
  ))

# add variable for source
df <- df %>%
  mutate(source_prev_rrt = case_when(ukrr_2019==1 & 
                                      latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "UPS",
                                      ukrr_2019!=1 & 
                                      latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "PS",
                                     ukrr_2019==1 & 
                                      !(latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown")) &
                                      latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "US",
                                     ukrr_2019==1 & 
                                      latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      !(latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown")) ~ "UP",
                                     ukrr_2019==1 & 
                                      !(latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown")) &
                                      !(latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown")) ~ "U",
                                     ukrr_2019!=1 & 
                                      latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      !(latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown")) ~ "P",
                                     ukrr_2019!=1 & 
                                      !(latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown")) &
                                      latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "S"))



# order factor
source_levels<-c("UPS", "UP", "US", "PS", "U", "P", "S")
df$source_prev_rrt<-factor(df$source_prev_rrt, levels=source_levels)


#calculate latest CKD stage in secondary care (the variable latest_ckd_status_secondary includes RRT which we don't want here)
#we have the dates of each ICD10 code for the CKD stages, so want to pick the latest one
#first get the dates of the ckd stages in secondary care

ckd_secondary_dates1<-df[,c("ip_ckd1_diagnosis_date", "ip_ckd2_diagnosis_date", "ip_ckd3_diagnosis_date",
                           "ip_ckd4_diagnosis_date", "ip_ckd5_diagnosis_date")]  %>%
  mutate_all( function(x) as.Date(x))  #convert to date

#vector of where there are no CKD codes at all
no_ckd<-rowSums(is.na(ckd_secondary_dates1))==5

#making changes needed to use max.col function
ckd_secondary_dates2<-ckd_secondary_dates1 %>%
         replace(is.na(.), as.Date("1900-01-01")) %>%  # replace missing with earliest date for max.col
          mutate_all(function(x) as.numeric(x))  #convert to numeric for max.col

#get which column has the latest date - the fist col is stage 1, second is stage 2 etc
ckd_stage_secondary<-max.col(ckd_secondary_dates2, ties.method="last")
#repalce with missing when all there is no ckd code (as earlier we replace the missings with an early date "1900-01-01")
ckd_stage_secondary[no_ckd]<-NA

#setting as a column
df$ckd_stage_sec<-ckd_stage_secondary


#apply disclosure control - redact if less than 7, then round to nearest 5
disclosure <-function(x, whichmargins=0){
  redact <- 0<x & x<=7  #flag redacted values for later
  x[redact] <- 0  # set redacted to 0
  x<-round(x/5)*5 # round to nearest 5
  if(whichmargins!=0){
    x<-addmargins(x,whichmargins) #add totals
  }
  if(whichmargins==1){
    redact<-rbind(redact, rep(FALSE,ncol(redact))) #add extra row to redaction, otherwise won't blank correctly
  }
  if(whichmargins==2){
    redact<-cbind(redact, rep(FALSE,nrow(redact))) #add extra col to redaction, otherwise won't blank correctly
  }
  x[redact]<-"<=7"  # flag redacted values
  return(x)
}

```

## Checking eGFR  
### eGFR category from Louis vs eGFR calculated on last value only
```{r, echo=FALSE, results='asis', error=TRUE}
k<- table(df$ckd_egfr_category, df$ckd_egfr_category_lastvalue)
k2<-k%>%
  kbl() %>%
  kable_styling()
print(k2)
```
### eGFR category assuming highest CKD stage when there is > or >= operator vs eGFR calculated on last value only
```{r, echo=FALSE, results='asis', error=TRUE}
k<- table(df$ckd_egfr_category_operator, df$ckd_egfr_category_lastvalue)
k2<-k%>%
  kbl() %>%
  kable_styling()
print(k2)
```


# 1a - Latest CKD codes in PC/SC care for people on KRT in PC/SC and not UKRR, by modality  
## Prevalent 1/1/2020
### in PC and not UKRR by PC modality
```{r, echo=FALSE, results='asis', error=TRUE}
k<-disclosure(table(df[df$source_prev_rrt %in% c("P", "PS"),]$latest_rrt_status,
      df[df$source_prev_rrt %in% c("P", "PS"),]$ckd_primis_stage)  ,1)
k2<-k%>%
  kbl() %>%
  add_header_above(c("KRT status", "CKD stage" = ncol(k))) %>%
  kable_styling()
print(k2)
```

### in SC and not UKRR by SC modality
```{r, echo=FALSE, results='asis', error=TRUE}
k<-disclosure(table(df[df$source_prev_rrt %in% c("S", "PS"),]$latest_rrt_status_secondary,
      df[df$source_prev_rrt %in% c("S", "PS"),]$ckd_stage_sec) ,1)
k2<-k%>%
  kbl() %>%
  add_header_above(c("KRT status", "CKD stage"  = ncol(k))) %>%
  kable_styling()
print(k2)
```

# 1b - For comparison, CKD codes in PC/SC care for people on KRT in PC/SC and also UKRR, by modality  
## Prevalent 1/1/2020
### in PC and UKRR by PC modality
```{r, echo=FALSE, results='asis', error=TRUE}
k<-disclosure(table(df[df$source_prev_rrt %in% c("UP", "UPS"),]$latest_rrt_status,
      df[df$source_prev_rrt %in% c("UP", "UPS"),]$ckd_primis_stage)  ,1)
k2<-k%>%
  kbl() %>%
  add_header_above(c("KRT status", "CKD stage" = ncol(k))) %>%
  kable_styling()
print(k2)
```

### in SC and UKRR, by SC modality
```{r, echo=FALSE, results='asis', error=TRUE}
k<-disclosure(table(df[df$source_prev_rrt %in% c("US", "UPS"),]$latest_rrt_status_secondary,
      df[df$source_prev_rrt %in% c("US", "UPS"),]$ckd_stage_sec) ,1)
k2<-k%>%
  kbl() %>%
  add_header_above(c("KRT status", "CKD stage"  = ncol(k))) %>%
  kable_styling()
print(k2)
```

# 2a - CKD staging by eGFR in PC for people on KRT in UKRR and not PC, by modality  
## Prevalent 1/1/2020
### in PC and not UKRR by PC modality
```{r, echo=FALSE, results='asis', error=TRUE}
k<-disclosure(table(df[df$source_prev_rrt %in% c("P", "PS"),]$latest_rrt_status,
      df[df$source_prev_rrt %in% c("P", "PS"),]$ckd_egfr_category)  ,1)
k2<-k%>%
  kbl() %>%
  add_header_above(c("KRT status", "CKD stage" = ncol(k))) %>%
  kable_styling()
print(k2)
```

# 2b - For comparison, CKD staging by eGFR in PC for people on KRT in PC and also UKRR, by modality  
## Prevalent 1/1/2020
### in PC and UKRR by PC modality
```{r, echo=FALSE, results='asis', error=TRUE}
k<-disclosure(table(df[df$source_prev_rrt %in% c("UP", "UPS"),]$latest_rrt_status,
      df[df$source_prev_rrt %in% c("UP", "UPS"),]$ckd_egfr_category)  ,1)
k2<-k%>%
  kbl() %>%
  add_header_above(c("KRT status", "CKD stage" = ncol(k))) %>%
  kable_styling()
print(k2)
```
# 3a - CKD staging by eGFR (last only) in PC for people on KRT in UKRR and not PC, by modality  
## Prevalent 1/1/2020
### in PC and not UKRR by PC modality
```{r, echo=FALSE, results='asis', error=TRUE}
k<-disclosure(table(df[df$source_prev_rrt %in% c("P", "PS"),]$latest_rrt_status,
      df[df$source_prev_rrt %in% c("P", "PS"),]$ckd_egfr_category_lastvalue)  ,1)
k2<-k%>%
  kbl() %>%
  add_header_above(c("KRT status", "CKD stage" = ncol(k))) %>%
  kable_styling()
print(k2)
```

# 3b - For comparison, CKD staging by eGFR (last only) in PC for people on KRT in PC and also UKRR, by modality  
## Prevalent 1/1/2020
### in PC and UKRR by PC modality
```{r, echo=FALSE, results='asis', error=TRUE}
k<-disclosure(table(df[df$source_prev_rrt %in% c("UP", "UPS"),]$latest_rrt_status,
      df[df$source_prev_rrt %in% c("UP", "UPS"),]$ckd_egfr_category_lastvalue)  ,1)
k2<-k%>%
  kbl() %>%
  add_header_above(c("KRT status", "CKD stage" = ncol(k))) %>%
  kable_styling()
print(k2)
```
 
# 4a - CKD staging by eGFR (highest stage in presence of > or >= operator) in PC for people on KRT in UKRR and not PC, by modality  
## Prevalent 1/1/2020
### in PC and not UKRR by PC modality
```{r, echo=FALSE, results='asis', error=TRUE}
k<-disclosure(table(df[df$source_prev_rrt %in% c("P", "PS"),]$latest_rrt_status,
      df[df$source_prev_rrt %in% c("P", "PS"),]$ckd_egfr_category_operator)  ,1)
k2<-k%>%
  kbl() %>%
  add_header_above(c("KRT status", "CKD stage" = ncol(k))) %>%
  kable_styling()
print(k2)
```

# 4b - For comparison, CKD staging by eGFR  (highest stage in presence of > or >= operator) in PC for people on KRT in PC and also UKRR, by modality  
## Prevalent 1/1/2020
### in PC and UKRR by PC modality
```{r, echo=FALSE, results='asis', error=TRUE}
k<-disclosure(table(df[df$source_prev_rrt %in% c("UP", "UPS"),]$latest_rrt_status,
      df[df$source_prev_rrt %in% c("UP", "UPS"),]$ckd_egfr_category_operator)  ,1)
k2<-k%>%
  kbl() %>%
  add_header_above(c("KRT status", "CKD stage" = ncol(k))) %>%
  kable_styling()
print(k2)
```
 
