---
title: "Analysis of incident patients"
author: "Shalini Santhakumaran"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

This document describes the cohorts of incident to renal replacement therapy during 2020 using 3 sources of data: the UK Renal Registry (UKRR), primary care data from TPP practices (PC) and secondary care inpatient and outpatient data (SC).  
Done separately to the rest of the analyis as we want to look at primary care data up to the end of 2020

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Load libraries
library(tidyverse)
library(knitr)
library(kableExtra)

# Create directory if it doesnt exist
fs::dir_create(here::here("output", "notebooks"))

demogs<-c("age_band", "sex", "ethnicity", "imd", "diabetes", "hypertension",  "at_risk", "region")

#read in dataset
# Considered picking up 1/1/2021 cohort but then we won't have anyone who died in 2020
#instead do a separate analysis of the nearest event  
 df<-read.csv(here::here("output", "ukrr", "input_ukrr_2020-01-01.csv.gz")) 

#getting year of start for incident start dates
inc_year_pc<-as.numeric(substr(df$incident_rrt_date,1,4))
inc_year_sc<-as.numeric(substr(df$incident_rrt_date_secondary,1,4))
#ukrr - start dates are stored in different variables, start with inc 2020 then look at prevalent cohorts, replacing missing values only
#will be consistent as we only have one start date per patient in the data that was sent
inc_year_ukrr<-as.numeric(substr(df$ukrr_inc2020_date,1,4))
inc_year_ukrr[is.na(inc_year_ukrr)]<-as.numeric(substr(df$ukrr_2019_startdate,1,4))
inc_year_ukrr[is.na(inc_year_ukrr)]<-as.numeric(substr(df$ukrr_2020_startdate,1,4))
inc_year_ukrr[is.na(inc_year_ukrr)]<-as.numeric(substr(df$ukrr_2021_startdate,1,4))
 

#add variable for dialysis/tx for UKRR
df<- df %>%
  mutate(ukrr_dtx_inc = ifelse(ukrr_inc2020_mod %in% c("ICHD", "HD", "HHD", "PD") , "Dialysis", ifelse(ukrr_inc2020_mod == "Tx", "Transplant", "Not on RRT"))) 
# order factor
df$ukrr_dtx_inc<-factor(df$ukrr_dtx_inc, levels=c("Not on RRT", "Dialysis", "Transplant"))


#read in codelists so we can tabulate the names
dialysis_codelist<-unique(read.csv(here::here("codelists", "opensafely-dialysis.csv"))[,1:2])
kidney_tx_codelist<-unique(read.csv(here::here("codelists", "opensafely-kidney-transplant.csv"))[,1:2])
all.codelists<-list(dialysis_codelist, kidney_tx_codelist )


# function to filter dataset to exclude patients with no RRT in either OS or UKRR as this is what we want to show pct by
keepRRT<-function(x){
  filter(x , ukrr_prev==1 | latest_renal_status %in% c("Dialysis", "Transplant", "RRT_unknown"))
}
keepRRT_nockd<-function(x){
  filter(x , ukrr_prev==1 | latest_rrt_status %in% c("Dialysis", "Transplant", "RRT_unknown"))
}

#apply disclosure control to a table - redact if less than 7, then round to nearest 5
disclosure <-function(x, whichmargins=0){
  redact <- 0<x & x<=7  #flag redacted values for later
  x[redact] <- 0  # set redacted to 0
  x<-round(x/5)*5 # round to nearest 5
  if(whichmargins!=0){
    x<-addmargins(x,whichmargins) #add totals
  }
  if(whichmargins==1){
    redact<-rbind(redact, rep(FALSE,ncol(redact))) #add extra row to redaction, otherwise won't blank correctly
  }
  if(whichmargins==2){
    redact<-cbind(redact, rep(FALSE,nrow(redact))) #add extra col to redaction, otherwise won't blank correctly
  }
  x[redact]<-"<=7"  # flag redacted values
  return(x)
}

```

# Section 1 - Compare incident start dates in PC, SC and UKRR
Use this output to look at PC/SC start dates for people starting in 2020 in the UKRR, and for UKRR start dates for people starting in PC/SC in 2020.

```{r, echo=FALSE, results='asis', error=TRUE}
  
t<-disclosure(table(inc_year_sc, inc_year_ukrr, useNA="ifany"))
  k<- t %>%
          kbl(caption="SC start date and UKRR start date", booktabs=TRUE) %>%
          add_header_above(c("Inc year SC", "Inc year UKRR"=ncol(t))) %>%
          kable_styling()
  print(k)
  
t<-disclosure(table(inc_year_pc, inc_year_ukrr, useNA="ifany"))
  k<- t %>%
          kbl(caption="PC start date and UKRR start date", booktabs=TRUE) %>%
          add_header_above(c("Inc year PC", "Inc year UKRR"=ncol(t))) %>%
          kable_styling()
  print(k)

t<-disclosure(table(inc_year_pc, inc_year_sc, useNA="ifany"))
  k<- t %>%
          kbl(caption="PC start date and UKRR start date", booktabs=TRUE) %>%
          add_header_above(c("Inc year PC", "Inc year SC"=ncol(t))) %>%
          kable_styling()
  print(k)

```

# Section 2 - Look at CKD status at the start of 2020 for people starting in 2020 in UKRR who are not incident in PC/SC at all

```{r, echo=FALSE, results='asis', error=TRUE}
# note that in this Rmd file incident_rrt_status refers to the incident status whenever they started, not just in 2020 like in other programs
k<-disclosure(table(df[df$ukrr_inc2020==1 & df$incident_rrt_status=="None",]$ukrr_dtx_inc,
      df[df$ukrr_inc2020==1 & df$incident_rrt_status=="None",]$latest_renal_status),0)
k2<-k%>%
  kbl(caption="Latest CKD status at start of 2020 for people not incident in PC at all") %>%
  add_header_above(c("UKRR", "Primary care" = ncol(k))) %>%
  kable_styling()
print(k2)


k<-disclosure(table(df[df$ukrr_inc2020==1  & df$incident_rrt_status_secondary=="None",]$ukrr_dtx_inc,
      df[df$ukrr_inc2020==1  & df$incident_rrt_status_secondary=="None",]$latest_ckd_status_secondary),0) 
k2<-k %>%
  kbl(caption="Latest CKD status at start of 2020 for people not incident in SC at all") %>%
  add_header_above(c("UKRR", "Secondary care" = ncol(k))) %>%
  kable_styling()
print(k2)

```



