---
title: "Results for Kidney Week presentation"
author: "Shalini Santhakumaran"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

This report is a subset of analysis from "Investigating discrepancies between UKRR and OpenSAFELY" (ukrr-discrepancies) with output for UK Renal Registry presentation for Kidney Week.

Cells less than or equal to 5 are shown as <=5 and corresponding percentages blanked. Medians are rounded to the nearest 5. All other numbers are rounded up to the nearest 5 to reduce risk of disclosure from subsequent versions, and percentages calculated on the rounded number.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Load libraries
library(tidyverse)
library(knitr)
library(kableExtra)
library(here)

# Create directory if it doesnt exist
fs::dir_create(here::here("output", "notebooks"))

#set up empty list of datasets and demog variables
dflist<-list()
years<-c("2019", "2020", "2021")
demogs<-c("sex", "age_band", "region", "imd", "diabetes", "at_risk", "hypertension")
for (i in 1:3){
  #read in dataset
  
#pick up 1st Jan cohort as latest renal status is on or before index date
 df<-read.csv(here::here("output", "joined", paste0("input_", as.numeric(years[i])+1,"-01-01.csv.gz"))) 

  #set prevalent data flag and modality
  df$ukrr_prev<-as.vector(t(df[paste0("ukrr_", years[i])]))
  df$ukrr_mod<-as.vector(t(df[paste0("ukrr_", years[i],"_mod")]))
  #add variable for dialysis/tx for UKRR
  df<- df %>%
    mutate(ukrr_dtx = ifelse(ukrr_mod %in% c("ICHD", "HD", "HHD", "PD") , "Dialysis", ifelse(ukrr_mod == "Tx", "Transplant", "Not on RRT")))
  # add variable for dialysis/tx for primary care
  df<- df %>%
    mutate(latest_dtx = ifelse(latest_renal_status %in% c("Dialysis", "Transplant", "RRT_unknown") , latest_renal_status, "Not on RRT"))
  # order factor
  df$ukrr_dtx<-factor(df$ukrr_dtx, levels=c("Not on RRT", "Dialysis", "Transplant"))
  df$latest_dtx<-factor(df$latest_dtx, levels=c("Not on RRT", "Dialysis", "Transplant", "RRT_unknown"))
  # add variable for source
  df<-df %>%
    mutate(rrt_source = ifelse(ukrr_prev==1 & latest_renal_status %in%  c("Dialysis", "Transplant", "RRT_unknown"), "Both",
                               ifelse (ukrr_prev==1 & !(latest_renal_status %in%  c("Dialysis", "Transplant", "RRT_unknown")), "UKRR only",
                                    ifelse(ukrr_prev!=1 & latest_renal_status %in%  c("Dialysis", "Transplant", "RRT_unknown"), "PC only",
                                           "Neither")))) %>%
    mutate(region = replace(region, region=="", "Missing"))
  # pick out latest code
  df<-df %>%
    mutate(latest_renal_code = ifelse(latest_renal_status=="None", "None",
                                  ifelse(latest_renal_status=="Dialysis", dialysis_code,
                                      ifelse(latest_renal_status=="Transplant", kidney_tx_code,
                                             ifelse(latest_renal_status=="RRT_unknown", RRT_code,
                                                  ifelse(latest_renal_status %in% paste0("CKD", seq(5)), ckd_primis_1_5_code,
                                                    ifelse(latest_renal_status == "CKD_unknown", ckd_code, "Uncategorised")))))))
  dflist[[i]] <- df
}

# function to filter dataset to exclude patients with no RRT in either OS or UKRR as this is what we want to show pct by
keepRRT<-function(x){
  filter(x , ukrr_prev==1 | latest_renal_status %in% c("Dialysis", "Transplant", "RRT_unknown"))
}

#function to round up to nearest accuracy
roundup = function(x, accuracy, f=ceiling){f(x/ accuracy) * accuracy}

#read in codelists so we can tabulate the names
ckd_codelist<-unique(read.csv(here::here("codelists", "opensafely-chronic-kidney-disease-snomed.csv"))[,1:2])
ckd_primis_1_5_codelist<-unique(read.csv(here::here("codelists", "user-Louis-ckd-stage.csv"))[,1:2])
dialysis_codelist<-unique(read.csv(here::here("codelists", "opensafely-dialysis.csv"))[,1:2])
kidney_tx_codelist<-unique(read.csv(here::here("codelists", "opensafely-kidney-transplant.csv"))[,1:2])
RRT_codelist<-unique(read.csv(here::here("codelists", "opensafely-renal-replacement-therapy.csv"))[,1:2])
all.codelists<-list(ckd_codelist, ckd_primis_1_5_codelist, dialysis_codelist, kidney_tx_codelist, RRT_codelist )
all.code.vars<-c("ckd_code", "ckd_primis_1_5_code", "dialysis_code", "kidney_tx_code", "RRT_code")

```

## 1 People on RRT according to the UKRR but not primary care data
First looking at at the renal status as defined by the latest primary care code.
```{r, echo=FALSE, results='asis'}
#First tabulate the latest renal status
list.of.dts<-lapply(dflist, function (x) 
  x %>%
    filter(rrt_source=="UKRR only") %>%
    count(ukrr_dtx, latest_renal_status) %>%
    spread(latest_renal_status, n, fill=0) %>%
    mutate(across(where(is.numeric)&!c(1), ~ roundup(.x,5))) %>% 
    mutate(across(where(is.numeric)&!c(1),~ifelse(.<=5,"<=5",.)))) 

for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=paste0("Table 1.1",i, " Latest PC renal status for people on RRT in the UKRR but not PC ",  years[i]), booktabs=TRUE) %>%
          #add_header_above(c("", "Primary care" = ncol(list.of.dts[[i]]) - 1)) %>%
          kable_styling()
  print(k)
}

```


## 2 People on RRT according to primary care data but not the UKRR
Below we look at what the most recent primary care codes are for these patients, by modality, and the median time since that code
```{r, echo=FALSE, results='asis'}

# Now plotting the time between last RRT code and prevalent date
plot.df<-mapply(function(x,y)
   x %>%
        filter(rrt_source=="PC only") %>%
        mutate(days_since_last_RRT = as.numeric(as.Date(paste0("1jan",as.numeric(y)+1), "%d%b%Y") - as.Date(latest_renal_date, "%Y-%m-%d"))),
   dflist, years, SIMPLIFY = FALSE)

# Looking at median time since code
codes.table.3<-function(df,status, codelist){
  df %>%
    filter(rrt_source=="PC only" & latest_renal_status==status) %>%
    group_by(latest_renal_code)%>%
    summarise(n=n(), median=median(days_since_last_RRT)) %>%
    left_join(codelist, by=setNames(names(codelist)[1], "latest_renal_code")) %>%
    mutate(latest_renal_code= replace(latest_renal_code, latest_renal_code %in% c("", NA) , "Missing")) %>%
      mutate(n=ifelse(n<=5,0,n)) %>% #first replace n<=5 with 0
      mutate(n=roundup(n,5)) %>%  #round up to nearest 5
      mutate(Percentage = ifelse(n==0, NA, round(100*n/sum(n), 0))) %>%   #then calculate %s, blanking if n is 0
        mutate(median = ifelse(n==0, NA, round(median/5)*5)) %>%   #then calculate %s, blanking if n is 0
        arrange(desc(n))  %>% #descending sort by n
      mutate(n=ifelse(n<=5,"<=5",n)) #now replacing <=5 (which is now 0 anyway) with "<=5"
}

list.of.dts<-lapply(plot.df, function(x) 
  mapply(function(y, z) 
    codes.table.3(x, y, z), 
    c("Dialysis", "Transplant"),  all.codelists[3:4], SIMPLIFY=FALSE))

for(i in 1:length(list.of.dts)) {
  for(j in 1:length(list.of.dts[[i]])){
  # names<-unique(colnames(list.of.dts[[i]][[j]])) #save existing colnames
  #  colnames(list.of.dts[[i]][[j]])<-rep("",length(names)*2) # blank colnames
   k<- list.of.dts[[i]][[j]]%>%
          kbl(caption=paste0("Table 3.", i, "." , j, 
                             " Median time since  ", 
                             c("Dialysis", "Transplant")[j], 
                             ", ", 
                             years[i] ), booktabs=TRUE) %>% 
          #add_header_above(c(demogs[j], rep(c("n", "%"), length(names)))) %>%
          #add_header_above(data.frame(c("", names), c(1,rep(2, length(names))))) %>%  #setting colnames as a dataframe 
          kable_styling()
  print(k)     
  }
}

```
## 4 Comparison table of time since last code for patients where UKRR and primary data agree
```{r, echo=FALSE, results='asis'}

comparison.df<-mapply(function(x,y)
   x %>%
        mutate(days_since_last_RRT = as.numeric(as.Date(paste0("1jan",as.numeric(y)+1), "%d%b%Y") - as.Date(latest_renal_date, "%Y-%m-%d"))),
   dflist, years, SIMPLIFY = FALSE)


# Looking at median time since code
codes.table.4<-function(df,status, codelist){
  df %>%
    filter(rrt_source=="Both" & as.character(latest_dtx)==as.character(ukrr_dtx) & latest_renal_status==status) %>%
    group_by(latest_renal_code)%>%
    summarise(n=n(), median=median(days_since_last_RRT)) %>%
    left_join(codelist, by=setNames(names(codelist)[1], "latest_renal_code")) %>%
    mutate(latest_renal_code= replace(latest_renal_code, latest_renal_code %in% c("", NA) , "Missing")) %>%
      mutate(n=ifelse(n<=5,0,n)) %>% #first replace n<=5 with 0
      mutate(n=roundup(n,5)) %>%  #round up to nearest 5
      mutate(Percentage = ifelse(n==0, NA, round(100*n/sum(n), 0))) %>%   #then calculate %s, blanking if n is 0
        mutate(median = ifelse(n==0, NA, round(median/5)*5)) %>%   #then calculate %s, blanking if n is 0
        arrange(desc(n))  %>% #descending sort by n
      mutate(n=ifelse(n<=5,"<=5",n)) #now replacing <=5 (which is now 0 anyway) with "<=5"
}

list.of.dts<-lapply(comparison.df, function(x) 
  mapply(function(y, z) 
    codes.table.4(x, y, z), 
    c("Dialysis", "Transplant"),  all.codelists[3:4], SIMPLIFY=FALSE))

for(i in 1:length(list.of.dts)) {
  for(j in 1:length(list.of.dts[[i]])){
  # names<-unique(colnames(list.of.dts[[i]][[j]])) #save existing colnames
  #  colnames(list.of.dts[[i]][[j]])<-rep("",length(names)*2) # blank colnames
   k<- list.of.dts[[i]][[j]]%>%
          kbl(caption=paste0("Table 4.", i, "." , j, 
                             " Median time since  ", 
                             c("Dialysis", "Transplant")[j], 
                             ", ", 
                             years[i] ), booktabs=TRUE) %>% 
          #add_header_above(c(demogs[j], rep(c("n", "%"), length(names)))) %>%
          #add_header_above(data.frame(c("", names), c(1,rep(2, length(names))))) %>%  #setting colnames as a dataframe 
          kable_styling()
  print(k)     
  }
}

```
