---
title: "Modality difference table"
author: "Shalini Santhakumaran"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

This document describes the cohorts of patients prevalent to renal replacement therapy (RRT) at 1st January 2020, and incident to renal replacement therapy during 2020 using 3 sources of data: the UK Renal Registry (UKRR), primary care data from TPP practices (PC) and secondary care inpatient and outpatient data (SC).  
Extra table looking at modality differences for people in UKRR and PC/SC as there was an error in the code before.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Load libraries
library(tidyverse)
library(knitr)
library(kableExtra)

# Create directory if it doesnt exist
fs::dir_create(here::here("output", "notebooks"))

#read in dataset
#pick up 1st Jan cohort as latest renal status is on or before index date
 df<-read.csv(here::here("output", "joined", "input_2020-01-01.csv.gz")) 
 
#add variable for dialysis/tx for UKRR
df<- df %>%
  mutate(ukrr_dtx = ifelse(ukrr_2019_mod %in% c("ICHD", "HD", "HHD", "PD") , "Dialysis", ifelse(ukrr_2019_mod == "Tx", "Transplant", "Not on RRT")))

# order factor
df$ukrr_dtx<-factor(df$ukrr_dtx, levels=c("Not on RRT", "Dialysis", "Transplant"))

# add variable for source
df <- df %>%
  mutate(source_prev_rrt = case_when(ukrr_2019==1 & 
                                      latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "UPS",
                                      ukrr_2019!=1 & 
                                      latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "PS",
                                     ukrr_2019==1 & 
                                      !(latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown")) &
                                      latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "US",
                                     ukrr_2019==1 & 
                                      latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      !(latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown")) ~ "UP",
                                     ukrr_2019==1 & 
                                      !(latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown")) &
                                      !(latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown")) ~ "U",
                                     ukrr_2019!=1 & 
                                      latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      !(latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown")) ~ "P",
                                     ukrr_2019!=1 & 
                                      !(latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown")) &
                                      latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "S")) 
# order factor
source_levels<-c("UPS", "UP", "US", "PS", "U", "P", "S")
df$source_prev_rrt<-factor(df$source_prev_rrt, levels=source_levels)


#function to round up to nearest accuracy
roundup = function(x, accuracy, f=ceiling){f(x/ accuracy) * accuracy}
```



Of people who are captured in UKRR and PC/SC, do the modalitie agree?   

## Prevalent 1/1/2020
### UKRR and PC
```{r, echo=FALSE, results='asis', error=TRUE}
k<-roundup(table(df[df$source_prev_rrt %in% c("UP", "UPS"),]$ukrr_dtx,
      df[df$source_prev_rrt %in% c("UP", "UPS"),]$latest_rrt_status),5)
k2<-k%>%
  kbl() %>%
  add_header_above(c("UKRR", "Primary care" = 3)) %>%
  kable_styling()
print(k2)
```
### UKRR and SC
```{r, echo=FALSE, results='asis', error=TRUE}
k<-roundup(table(df[df$source_prev_rrt %in% c("US", "UPS"),]$ukrr_dtx,
      df[df$source_prev_rrt %in% c("US", "UPS"),]$latest_rrt_status_secondary),5) 
k2<-k%>%
  kbl() %>%
  add_header_above(c("UKRR", "Secondary care" = ncol(k))) %>%
  kable_styling()
print(k2)
```

