---
title: "Data checks for UKRR OS comparison"
author: "Shalini Santhakumaran"
date: '2022-04-24'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
# Create directory if it doesnt exist
fs::dir_create(here::here("output", "notebooks"))
#read in - just doing 2019 as only to check variables are being derived correctly
d<-read.csv(here::here("output", "input_2019-12-01.csv.gz"))
ukrr<-read.csv(here::here("output", "input_ukrr_ethnicity.csv.gz"))
#redoing join so we get a left join with the UKRR data on the left - the cohort joiner has the monthly file on the left
#still might not get demogs for all patients as only adults alive and registered at start of December 2019
d1<-merge(x=ukrr, y=d, by="patient_id", all.x=TRUE)

#set up
ukrr_cohorts<-d1[c("ukrr_2019", "ukrr_2020", "ukrr_2021", "ukrr_inc2020")]
ukrr_mod<-d1[c("ukrr_2019_mod", "ukrr_2020_mod", "ukrr_2021_mod", "ukrr_inc2020_mod")]
ukrr_centre<-d1[c("ukrr_2019_centre", "ukrr_2020_centre", "ukrr_2021_centre", "ukrr_inc2020_centre")]
demogs<-d1[c("sex", "age_band", "region", "imd", "diabetes", "at_risk", "hypertension")]
demogs <- lapply(demogs, as.factor)
```
## Latest renal status

First checking that the latest renal status is being determined correctly in the study definition (relationships defined using patients.categorised_as don't hold, the variables are are just created using dummy data expectations)

Binary flags for primary care codelists where latest_renal_status = "None" (i.e. the patient is classed as never having had any stage of CKD)
```{r, echo=FALSE}
kable(apply(d1[which(d1$latest_renal_status=="None"),
         c("dialysis","kidney_tx","RRT","ckd_primis_1_5")],
        2,table ))
```

Checking latest renal status for patients whose last date was dialysis
```{r, echo=FALSE}
kable(d1[which(d1$dialysis_date==d1$latest_renal_date & 
              d1$kidney_tx_date!=d1$latest_renal_date),]%>%
  count(latest_renal_status))
```
Checking latest renal status for patients whose last date was tx
```{r, echo=FALSE}
kable(d1[which(d1$dialysis_date!=d1$latest_renal_date & 
              d1$kidney_tx_date==d1$latest_renal_date),]%>%
  count(latest_renal_status))
```
Checking latest renal status for patients whose last date was RRT but none of the above apply
```{r, echo=FALSE}
kable(d1[which(d1$dialysis_date==d1$latest_renal_date |
                d1$kidney_tx_date==d1$latest_renal_date |
                d1$RRT_date==d1$latest_renal_date),]%>%
  count(latest_renal_status))
```
Checking latest renal status for patients whose last date was CKD
```{r, echo=FALSE}
kable(d1[which(d1$ckd_primis_1_5_date==d1$latest_renal_date),]%>%
  count(latest_renal_status))
```
Checking CKD stages
```{r, echo=FALSE}
d1[which(d1$latest_renal_date==d1$ckd_primis_1_5_date |
          d1$latest_renal_date==d1$ckd_date),]%>%
    group_by(latest_renal_status, ckd_primis_stage) %>%
    tally()%>%
    spread(latest_renal_status, n) %>%
    kable()
```
Checking uncategorised
```{r, echo=FALSE}
kable(apply(d1[which(d1$latest_renal_status=="Uncategorised"),
         c("dialysis","kidney_tx","RRT","ckd_primis_stage")],
      2,table ))
```
Checking demographics
```{r, echo=FALSE, results='asis'}
dtable <- function(x){
    df1<-table(x, d1$latest_renal_status)  #numbers
    df2<-prop.table(table(x,d1$latest_renal_status), margin=2) #proportions
    df3<-cbind(df1, df2) #combine
    colnames(df3)<-c(paste(colnames(df1), "N"), paste(colnames(df2), "%")) #name columns
    df4 <- df3[, c(matrix(1:ncol(df3), nrow = 2, byrow = T))]  
    return(df4)
}
list.of.dts<-lapply(demogs, dtable)
for(i in 1:length(list.of.dts)) {
  print(kable(x = list.of.dts[[i]], caption=paste0(names(demogs)[i], " by latest renal status")))
}
```


## Checking frequencies of UKRR daa
Modality by cohort
```{r,  echo=FALSE}
mapply(function(x, y) 
    rbind("N"=table(y[x==1]),
          "%"=prop.table(table(y[x==1]))) , 
    ukrr_cohorts, ukrr_mod, SIMPLIFY=FALSE)
```
Centre by cohort
```{r ,  echo=FALSE}
mapply(function(x, y) 
    cbind("N"=table(y[x==1]),
          "%"=prop.table(table(y[x==1]))) , 
    ukrr_cohorts, ukrr_centre, SIMPLIFY=FALSE)
```
Cohort demographics
```{r, echo=FALSE, results='asis'}
#tabulate each cohort by each demog
list.of.dts<-(lapply(demogs, function(y) 
  as.data.frame(lapply(ukrr_cohorts, function(x) 
    cbind("N"=table(y[x==1]),
          "%"=prop.table(table(y[x==1]))) ))))
for(i in 1:length(list.of.dts)) {
  print(kable(x = list.of.dts[[i]], caption=paste0(names(demogs)[i], " by UKRR cohort")))
}
```
