---
title: "Data for tables and figures - other years"
author: "Shalini Santhakumaran"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

This document describes the cohorts of patients prevalent to renal replacement therapy (RRT) at 1st January 2021, and incident to renal replacement therapy during 2021 using 3 sources of data: the UK Renal Registry (UKRR), primary care data from TPP practices (PC) and secondary care inpatient and outpatient data (SC).  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Load libraries
library(tidyverse)
library(knitr)
library(kableExtra)

# Create directory if it doesnt exist
fs::dir_create(here::here("output", "notebooks"))

demogs<-c("age_band", "sex", "ethnicity", "imd", "diabetes", "hypertension",  "at_risk", "region")

#read in dataset
#pick up 1st Jan cohort as latest renal status is on or before index date
 df<-read.csv(here::here("output", "ukrr", "input_ukrr_short_2021-01-01.csv.gz")) 
 

#add variable for dialysis/tx for UKRR
df<- df %>%
  mutate(ukrr_dtx = ifelse(ukrr_2020_mod %in% c("ICHD", "HD", "HHD", "PD") , "Dialysis", ifelse(ukrr_2020_mod == "Tx", "Transplant", "Not on RRT"))) %>%
  mutate(region = replace(region, region=="", "Missing"))
# order factor
df$ukrr_dtx<-factor(df$ukrr_dtx, levels=c("Not on RRT", "Dialysis", "Transplant"))

# add variable for source
df <- df %>%
  mutate(source_prev_rrt = case_when(ukrr_2020==1 & 
                                      latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "UPS",
                                      ukrr_2020!=1 & 
                                      latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "PS",
                                     ukrr_2020==1 & 
                                      !(latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown")) &
                                      latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "US",
                                     ukrr_2020==1 & 
                                      latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      !(latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown")) ~ "UP",
                                     ukrr_2020==1 & 
                                      !(latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown")) &
                                      !(latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown")) ~ "U",
                                     ukrr_2020!=1 & 
                                      latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      !(latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown")) ~ "P",
                                     ukrr_2020!=1 & 
                                      !(latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown")) &
                                      latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "S")) %>%
  mutate(source_prev_dialysis = case_when(ukrr_dtx=="Dialysis" & 
                                      latest_rrt_status  =="Dialysis" &
                                      latest_rrt_status_secondary =="Dialysis" ~ "UPS",
                                      ukrr_dtx!="Dialysis" & 
                                      latest_rrt_status  =="Dialysis" &
                                      latest_rrt_status_secondary =="Dialysis" ~ "PS",
                                     ukrr_dtx=="Dialysis" & 
                                      !(latest_rrt_status  =="Dialysis") &
                                      latest_rrt_status_secondary =="Dialysis" ~ "US",
                                     ukrr_dtx=="Dialysis" & 
                                      latest_rrt_status  =="Dialysis" &
                                      !(latest_rrt_status_secondary =="Dialysis") ~ "UP",
                                     ukrr_dtx=="Dialysis" & 
                                      !(latest_rrt_status  =="Dialysis") &
                                      !(latest_rrt_status_secondary =="Dialysis") ~ "U",
                                     ukrr_dtx!="Dialysis" & 
                                      latest_rrt_status  =="Dialysis" &
                                      !(latest_rrt_status_secondary =="Dialysis") ~ "P",
                                     ukrr_dtx!="Dialysis" & 
                                      !(latest_rrt_status  =="Dialysis") &
                                      latest_rrt_status_secondary =="Dialysis" ~ "S")) %>%  
    mutate(source_prev_transplant = case_when(ukrr_dtx=="Transplant" & 
                                      latest_rrt_status  =="Transplant" &
                                      latest_rrt_status_secondary =="Transplant" ~ "UPS",
                                      ukrr_dtx!="Transplant" & 
                                      latest_rrt_status  =="Transplant" &
                                      latest_rrt_status_secondary =="Transplant" ~ "PS",
                                     ukrr_dtx=="Transplant" & 
                                      !(latest_rrt_status  =="Transplant") &
                                      latest_rrt_status_secondary =="Transplant" ~ "US",
                                     ukrr_dtx=="Transplant" & 
                                      latest_rrt_status  =="Transplant" &
                                      !(latest_rrt_status_secondary =="Transplant") ~ "UP",
                                     ukrr_dtx == "Transplant" & 
                                      !(latest_rrt_status  =="Transplant") &
                                      !(latest_rrt_status_secondary =="Transplant") ~ "U",
                                     ukrr_dtx!="Transplant" & 
                                      latest_rrt_status  =="Transplant" &
                                      !(latest_rrt_status_secondary =="Transplant") ~ "P",
                                     ukrr_dtx!="Transplant" & 
                                      !(latest_rrt_status  =="Transplant") &
                                      latest_rrt_status_secondary =="Transplant" ~ "S")) 
# pick out latest RRT code for primary care
df<-df %>%
    mutate(latest_rrt_code = ifelse(latest_rrt_status=="None", "None",
                                  ifelse(latest_rrt_status=="Dialysis", dialysis_code,
                                      ifelse(latest_rrt_status=="Transplant", kidney_tx_code,
                                             ifelse(latest_rrt_status=="RRT_unknown", RRT_code,"Uncategorised")))))


# order factor
source_levels<-c("UPS", "UP", "US", "PS", "U", "P", "S")
df$source_prev_rrt<-factor(df$source_prev_rrt, levels=source_levels)
df$source_prev_dialysis<-factor(df$source_prev_dialysis, levels=source_levels)
df$source_prev_transplant<-factor(df$source_prev_transplant, levels=source_levels)

#read in codelists so we can tabulate the names
dialysis_codelist<-unique(read.csv(here::here("codelists", "opensafely-dialysis.csv"))[,1:2])
kidney_tx_codelist<-unique(read.csv(here::here("codelists", "opensafely-kidney-transplant.csv"))[,1:2])
all.codelists<-list(dialysis_codelist, kidney_tx_codelist )


#apply disclosure control - redact if less than 7, then round to nearest 5
disclosure <-function(x, whichmargins=0){
  redact <- 0<x & x<=7  #flag redacted values for later
  x[redact] <- 0  # set redacted to 0
  x<-round(x/5)*5 # round to nearest 5
  if(whichmargins!=0){
    x<-addmargins(x,whichmargins) #add totals
  }
  if(whichmargins==1){
    redact<-rbind(redact, rep(FALSE,ncol(redact))) #add extra row to redaction, otherwise won't blank correctly
  }
  if(whichmargins==2){
    redact<-cbind(redact, rep(FALSE,nrow(redact))) #add extra col to redaction, otherwise won't blank correctly
  }
  x[redact]<-"<=7"  # flag redacted values
  return(x)
}

```

# Section 1 Chacteristics of the study population
##  Prevalent at 1/1/2021, by data source and modality
### UKRR
```{r, echo=FALSE, results='asis', error=TRUE}
#create tables
list.of.dts<-lapply(df[df$ukrr_2020==1, demogs],
    function(x) disclosure(table(x, df[df$ukrr_2020==1,]$ukrr_dtx, useNA="ifany"),2))
#printing
for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=paste("Prevalent UKRR", demogs[i]), booktabs=TRUE) %>%
          kable_styling()
  print(k)
}
```
### Primary care
```{r, echo=FALSE, results='asis', error=TRUE}
#create tables
list.of.dts<-lapply(df[df$latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown"), demogs],
       function(x) disclosure(table(x, df[df$latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown"),]$latest_rrt_status, useNA="ifany"),2))
#printing
for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=paste("Prevalent primary care", demogs[i]), booktabs=TRUE) %>%
          kable_styling()
  print(k)
}
```
### Secondary care
```{r, echo=FALSE, results='asis', error=TRUE}
#create tables
list.of.dts<-lapply(df[df$latest_rrt_status_secondary  %in%  c("Dialysis", "Transplant", "RRT_unknown"), demogs],
       function(x) 
         disclosure(
           table(x, df[df$latest_rrt_status_secondary  %in%  c("Dialysis", "Transplant", "RRT_unknown"),]$latest_rrt_status_secondary, useNA="ifany")
           ,2))
#printing
for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=paste("Prevalent secondary care", demogs[i]), booktabs=TRUE) %>%
          kable_styling()
  print(k)
}
```


### add median age
```{r, echo=FALSE, results='asis', error=TRUE}
#create tables

med.age<-function(x){
  x <- enquo(x)
  df%>%
    group_by(!!x)%>% 
    summarise(age = quantile(age, c(0.25, 0.5, 0.75), na.rm=TRUE), q = c(0.25, 0.5, 0.75)) %>%
    spread(q, age)
  }

  k<- med.age(ukrr_dtx) %>%
          kbl(caption="Prevalent UKRR", booktabs=TRUE) %>%
          kable_styling()
  print(k)

    k<- med.age(latest_rrt_status ) %>%
          kbl(caption="Prevalent primary care", booktabs=TRUE) %>%
          kable_styling()
  print(k)
  
    k<- med.age(latest_rrt_status_secondary) %>%
          kbl(caption="Prevalent secondary care", booktabs=TRUE) %>%
          kable_styling()
  print(k)

```


# Section 2 Overall agreement  

Agreement is indicated by the data source, defined as UPS = in the **U**KRR, **P**rimary care and **S**econdary care data; UP = in **U**KRR and **P**rimary care only; U = in **U**KRR only etc.  

## Prevalent at 1/1/2021, overall and by modality
### All RRT
```{r, echo=FALSE, results='asis', error=TRUE}
disclosure(table(df$source_prev_rrt, useNA="ifany"),0) %>%
  kbl(col.names=c("Source", "N")) %>%
  kable_styling()
```
### Dialysis
```{r, echo=FALSE, results='asis', error=TRUE}
disclosure(table(df$source_prev_dialysis, useNA="ifany"),0) %>%
  kbl(col.names=c("Source", "N")) %>%
  kable_styling()
```
### Transplant
```{r, echo=FALSE, results='asis', error=TRUE}
disclosure(table(df$source_prev_transplant, useNA="ifany"),0) %>%
  kbl(col.names=c("Source", "N")) %>%
  kable_styling()
```

# Section 3 Agreement by demographic characteristics
## Prevalent at 1/1/2021, overall and by modality
### All RRT

```{r, echo=FALSE, results='asis', error=TRUE}
#create tables
list.of.dts<-lapply(df[demogs], function(x) disclosure(table(x, df$source_prev_rrt, useNA="ifany"),0))
#printing
for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=paste("Prevalent, all RRT", demogs[i]), booktabs=TRUE) %>%
          kable_styling()
  print(k)
}
```
### Dialysis
```{r, echo=FALSE, results='asis', error=TRUE}
#create tables
list.of.dts<-lapply(df[demogs], function(x) disclosure(table(x, df$source_prev_dialysis, useNA="ifany"),0))
#printing
for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=paste("Prevalent dialysis", demogs[i]), booktabs=TRUE) %>%
          kable_styling()
  print(k)
}
```
### Transplant
```{r, echo=FALSE, results='asis', error=TRUE}
#create tables
list.of.dts<-lapply(df[demogs], function(x) disclosure(table(x, df$source_prev_transplant, useNA="ifany"),0))
#printing
for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=paste("Prevalent transplant", demogs[i]), booktabs=TRUE) %>%
          kable_styling()
  print(k)
}
```

# Section 4  - Agreement of modality  

Of people who are captured in UKRR and PC/SC, do the modalities agree?   

## Prevalent 1/1/2021
### UKRR and PC
```{r, echo=FALSE, results='asis', error=TRUE}
k<-disclosure(table(df[df$source_prev_rrt %in% c("UP", "UPS"),]$ukrr_dtx,
      df[df$source_prev_rrt %in% c("UP", "UPS"),]$latest_rrt_status),0)
k2<-k%>%
  kbl() %>%
  add_header_above(c("UKRR", "Primary care" = 3)) %>%
  kable_styling()
print(k2)
```
### UKRR and SC
```{r, echo=FALSE, results='asis', error=TRUE}
k<-disclosure(table(df[df$source_prev_rrt %in% c("US", "UPS"),]$ukrr_dtx,
      df[df$source_prev_rrt %in% c("US", "UPS"),]$incident_rrt_status_secondary),0) 
k2<-k%>%
  kbl() %>%
  add_header_above(c("UKRR", "Secondary care" = ncol(k))) %>%
  kable_styling()
print(k2)
```


# Section 5 - CKD codes in PC/SC care for people in UKRR and not PC/SC, by modality  
## Prevalent 1/1/2021
### in UKRR and not PC
```{r, echo=FALSE, results='asis', error=TRUE}
k<-disclosure(table(df[df$source_prev_rrt %in% c("US", "U"),]$ukrr_dtx,
      df[df$source_prev_rrt %in% c("US", "U"),]$latest_renal_status, exclude="Not on RRT")  ,0)
k2<-k%>%
  kbl() %>%
  add_header_above(c("UKRR", "Primary care" = ncol(k))) %>%
  kable_styling()
print(k2)
```
### in UKRR and not SC
```{r, echo=FALSE, results='asis', error=TRUE}
k<-disclosure(table(df[df$source_prev_rrt %in% c("UP", "U"),]$ukrr_dtx,
      df[df$source_prev_rrt %in% c("UP", "U"),]$latest_ckd_status_secondary) ,0)
k2<-k%>%
  kbl() %>%
  add_header_above(c("UKRR", "Secondary care" = ncol(k))) %>%
  kable_styling()
print(k2)
```


 

