---
title: "Data for tables and figures"
author: "Shalini Santhakumaran"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Load libraries
library(tidyverse)
library(knitr)
library(kableExtra)

# Create directory if it doesnt exist
fs::dir_create(here::here("output", "notebooks"))

demogs<-c("age_band", "sex", "ethnicity", "imd", "diabetes", "hypertension",  "at_risk", "region")

#read in dataset
#pick up 1st Jan cohort as latest renal status is on or before index date
 df<-read.csv(here::here("output", "joined", "input_2020-01-01.csv.gz")) 
  
#add variable for dialysis/tx for UKRR
df<- df %>%
  mutate(ukrr_dtx = ifelse(ukrr_2019_mod %in% c("ICHD", "HD", "HHD", "PD") , "Dialysis", ifelse(ukrr_2019_mod == "Tx", "Transplant", "Not on RRT"))) %>%
  mutate(ukrr_dtx_inc = ifelse(ukrr_inc2020_mod %in% c("ICHD", "HD", "HHD", "PD") , "Dialysis", ifelse(ukrr_inc2020_mod == "Tx", "Transplant", "Not on RRT"))) %>%
  mutate(region = replace(region, region=="", "Missing"))
# order factor
df$ukrr_dtx<-factor(df$ukrr_dtx, levels=c("Not on RRT", "Dialysis", "Transplant"))
df$ukrr_dtx_inc<-factor(df$ukrr_dtx_inc, levels=c("Not on RRT", "Dialysis", "Transplant"))

# add variable for source
df <- df %>%
  mutate(source_prev_rrt = case_when(ukrr_2019==1 & 
                                      latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "UPS",
                                      ukrr_2019!=1 & 
                                      latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "PS",
                                     ukrr_2019==1 & 
                                      !(latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown")) &
                                      latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "US",
                                     ukrr_2019==1 & 
                                      latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      !(latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown")) ~ "UP",
                                     ukrr_2019==1 & 
                                      !(latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown")) &
                                      !(latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown")) ~ "U",
                                     ukrr_2019!=1 & 
                                      latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      !(latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown")) ~ "P",
                                     ukrr_2019!=1 & 
                                      !(latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown")) &
                                      latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "S")) %>%
  mutate(source_prev_dialysis = case_when(ukrr_dtx=="Dialysis" & 
                                      latest_rrt_status  =="Dialysis" &
                                      latest_rrt_status_secondary =="Dialysis" ~ "UPS",
                                      ukrr_dtx!="Dialysis" & 
                                      latest_rrt_status  =="Dialysis" &
                                      latest_rrt_status_secondary =="Dialysis" ~ "PS",
                                     ukrr_dtx=="Dialysis" & 
                                      !(latest_rrt_status  =="Dialysis") &
                                      latest_rrt_status_secondary =="Dialysis" ~ "US",
                                     ukrr_dtx=="Dialysis" & 
                                      latest_rrt_status  =="Dialysis" &
                                      !(latest_rrt_status_secondary =="Dialysis") ~ "UP",
                                     ukrr_dtx=="Dialysis" & 
                                      !(latest_rrt_status  =="Dialysis") &
                                      !(latest_rrt_status_secondary =="Dialysis") ~ "U",
                                     ukrr_dtx!="Dialysis" & 
                                      latest_rrt_status  =="Dialysis" &
                                      !(latest_rrt_status_secondary =="Dialysis") ~ "P",
                                     ukrr_dtx!="Dialysis" & 
                                      !(latest_rrt_status  =="Dialysis") &
                                      latest_rrt_status_secondary =="Dialysis" ~ "S")) %>%  
    mutate(source_prev_transplant = case_when(ukrr_dtx=="Transplant" & 
                                      latest_rrt_status  =="Transplant" &
                                      latest_rrt_status_secondary =="Transplant" ~ "UPS",
                                      ukrr_dtx!="Transplant" & 
                                      latest_rrt_status  =="Transplant" &
                                      latest_rrt_status_secondary =="Transplant" ~ "PS",
                                     ukrr_dtx=="Transplant" & 
                                      !(latest_rrt_status  =="Transplant") &
                                      latest_rrt_status_secondary =="Transplant" ~ "US",
                                     ukrr_dtx=="Transplant" & 
                                      latest_rrt_status  =="Transplant" &
                                      !(latest_rrt_status_secondary =="Transplant") ~ "UP",
                                     ukrr_dtx == "Transplant" & 
                                      !(latest_rrt_status  =="Transplant") &
                                      !(latest_rrt_status_secondary =="Transplant") ~ "U",
                                     ukrr_dtx!="Transplant" & 
                                      latest_rrt_status  =="Transplant" &
                                      !(latest_rrt_status_secondary =="Transplant") ~ "P",
                                     ukrr_dtx!="Transplant" & 
                                      !(latest_rrt_status  =="Transplant") &
                                      latest_rrt_status_secondary =="Transplant" ~ "S")) %>%  
    mutate(source_inc_rrt = case_when(ukrr_inc2020==1 & 
                                      incident_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      incident_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "UPS",
                                      ukrr_inc2020!=1 & 
                                      incident_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      incident_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "PS",
                                     ukrr_inc2020==1 & 
                                      !(incident_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown")) &
                                      incident_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "US",
                                     ukrr_inc2020==1 & 
                                      incident_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      !(incident_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown")) ~ "UP",
                                     ukrr_inc2020==1 & 
                                      !(incident_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown")) &
                                      !(incident_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown")) ~ "U",
                                     ukrr_inc2020!=1 & 
                                      incident_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      !(incident_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown")) ~ "P",
                                     ukrr_inc2020!=1 & 
                                      !(incident_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown")) &
                                      incident_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "S")) %>%
  mutate(source_inc_dialysis = case_when(ukrr_dtx_inc=="Dialysis" & 
                                      incident_rrt_status  =="Dialysis" &
                                      incident_rrt_status_secondary =="Dialysis" ~ "UPS",
                                      ukrr_dtx_inc!="Dialysis" & 
                                      incident_rrt_status  =="Dialysis" &
                                      incident_rrt_status_secondary =="Dialysis" ~ "PS",
                                     ukrr_dtx_inc=="Dialysis" & 
                                      !(incident_rrt_status  =="Dialysis") &
                                      incident_rrt_status_secondary =="Dialysis" ~ "US",
                                     ukrr_dtx_inc=="Dialysis" & 
                                      incident_rrt_status  =="Dialysis" &
                                      !(incident_rrt_status_secondary =="Dialysis") ~ "UP",
                                     ukrr_dtx_inc == "Dialysis" & 
                                      !(incident_rrt_status  =="Dialysis") &
                                      !(incident_rrt_status_secondary =="Dialysis") ~ "U",
                                     ukrr_dtx_inc!="Dialysis" & 
                                      incident_rrt_status  =="Dialysis" &
                                      !(incident_rrt_status_secondary =="Dialysis") ~ "P",
                                     ukrr_dtx_inc!="Dialysis" & 
                                      !(incident_rrt_status  =="Dialysis") &
                                      incident_rrt_status_secondary =="Dialysis" ~ "S")) %>%  
    mutate(source_inc_transplant = case_when(ukrr_dtx_inc=="Transplant" & 
                                      incident_rrt_status  =="Transplant" &
                                      incident_rrt_status_secondary =="Transplant" ~ "UPS",
                                      ukrr_dtx_inc!="Transplant" & 
                                      incident_rrt_status  =="Transplant" &
                                      incident_rrt_status_secondary =="Transplant" ~ "PS",
                                     ukrr_dtx_inc=="Transplant" & 
                                      !(incident_rrt_status  =="Transplant") &
                                      incident_rrt_status_secondary =="Transplant" ~ "US",
                                     ukrr_dtx_inc=="Transplant" & 
                                      incident_rrt_status  =="Transplant" &
                                      !(incident_rrt_status_secondary =="Transplant") ~ "UP",
                                     ukrr_dtx_inc == "Transplant" & 
                                      !(incident_rrt_status  =="Transplant") &
                                      !(incident_rrt_status_secondary =="Transplant") ~ "U",
                                     ukrr_dtx_inc!="Transplant" & 
                                      incident_rrt_status  =="Transplant" &
                                      !(incident_rrt_status_secondary =="Transplant") ~ "P",
                                     ukrr_dtx_inc!="Transplant" & 
                                      !(incident_rrt_status  =="Transplant") &
                                      incident_rrt_status_secondary =="Transplant" ~ "S"))
# order factor
source_levels<-c("UPS", "UP", "US", "PS", "U", "P", "S")
df$source_prev_rrt<-factor(df$source_prev_rrt, levels=source_levels)
df$source_prev_dialysis<-factor(df$source_prev_dialysis, levels=source_levels)
df$source_prev_transplant<-factor(df$source_prev_transplant, levels=source_levels)
df$source_inc_rrt<-factor(df$source_inc_rrt, levels=source_levels)
df$source_inc_dialysis<-factor(df$source_inc_dialysis, levels=source_levels)
df$source_inc_transplant<-factor(df$source_inc_transplant, levels=source_levels)


# function to filter dataset to exclude patients with no RRT in either OS or UKRR as this is what we want to show pct by
keepRRT<-function(x){
  filter(x , ukrr_prev==1 | latest_renal_status %in% c("Dialysis", "Transplant", "RRT_unknown"))
}
keepRRT_nockd<-function(x){
  filter(x , ukrr_prev==1 | latest_rrt_status %in% c("Dialysis", "Transplant", "RRT_unknown"))
}

#function to round up to nearest accuracy
roundup = function(x, accuracy, f=ceiling){f(x/ accuracy) * accuracy}
```

# Table 1 Chacteristics of the study population
## Prevalent at 1/1/2020
### UKRR
```{r, echo=FALSE, results='asis'}
#create tables
list.of.dts<-lapply(df[df$ukrr_2019==1, demogs],
    function(x) addmargins(table(x, df[df$ukrr_2019==1,]$ukrr_dtx, useNA="ifany"), 2))
#printing
for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=paste("Prevalent UKRR", demogs[i]), booktabs=TRUE) %>%
          kable_styling()
  print(k)
}
```
### Primary care 
```{r, echo=FALSE, results='asis'}
#create tables
list.of.dts<-lapply(df[df$latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown"), demogs],
       function(x) addmargins(table(x, df[df$latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown"),]$latest_rrt_status, useNA="ifany"), 2))
#printing
for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=paste("Prevalent primary care", demogs[i]), booktabs=TRUE) %>%
          kable_styling()
  print(k)
}
```
### Secondary care
```{r, echo=FALSE, results='asis'}
#create tables
list.of.dts<-lapply(df[df$latest_rrt_status_secondary  %in%  c("Dialysis", "Transplant", "RRT_unknown"), demogs],
       function(x) 
         addmargins(
           table(x, df[df$latest_rrt_status_secondary  %in%  c("Dialysis", "Transplant", "RRT_unknown"),]$latest_rrt_status_secondary, useNA="ifany")
           , 2))
#printing
for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=paste("Prevalent secondary care", demogs[i]), booktabs=TRUE) %>%
          kable_styling()
  print(k)
}
```

## Incident in 2020, by modality
### UKRR
```{r, echo=FALSE, results='asis'}
#create tables
list.of.dts<-lapply(df[df$ukrr_inc2020==1, demogs],
       function(x) addmargins(table(x, df[df$ukrr_inc2020==1,]$ukrr_dtx_inc, useNA="ifany"), 2))
#printing
for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=paste("Incident UKRR", demogs[i]), booktabs=TRUE) %>%
          kable_styling()
  print(k)
}
```
### Primary care 
```{r, echo=FALSE, results='asis'}
#create tables
list.of.dts<-lapply(df[df$incident_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown"), demogs],
       function(x) addmargins(table(x, df[df$incident_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown"),]$incident_rrt_status, useNA="ifany"), 2))
#printing
for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=paste("Incident primary care", demogs[i]), booktabs=TRUE) %>%
          kable_styling()
  print(k)
}
```
### Secondary care
```{r, echo=FALSE, results='asis'}
#create tables
list.of.dts<-lapply(df[df$incident_rrt_status_secondary  %in%  c("Dialysis", "Transplant", "RRT_unknown"), demogs],
       function(x) addmargins(
         table(x, df[df$incident_rrt_status_secondary  %in%  c("Dialysis", "Transplant", "RRT_unknown"),]$incident_rrt_status_secondary, useNA="ifany")
         , 2))
#printing
for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=paste("Incident secondary care", demogs[i]), booktabs=TRUE) %>%
          kable_styling()
  print(k)
}
```

# Table 2 Agreement by demographic characteristics
## Prevalent 1/1/2020
### All RRT

```{r, echo=FALSE, results='asis'}
#create tables
list.of.dts<-lapply(df[demogs], function(x) table(x, df$source_prev_rrt, useNA="ifany"))
#printing
for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=paste("Prevalent, all RRT", demogs[i]), booktabs=TRUE) %>%
          kable_styling()
  print(k)
}
```
### Dialysis
```{r, echo=FALSE, results='asis'}
#create tables
list.of.dts<-lapply(df[demogs], function(x) table(x, df$source_prev_dialysis, useNA="ifany"))
#printing
for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=paste("Prevalent dialysis", demogs[i]), booktabs=TRUE) %>%
          kable_styling()
  print(k)
}
```
### Transplant
```{r, echo=FALSE, results='asis'}
#create tables
list.of.dts<-lapply(df[demogs], function(x) table(x, df$source_prev_transplant, useNA="ifany"))
#printing
for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=paste("Prevalent transplant", demogs[i]), booktabs=TRUE) %>%
          kable_styling()
  print(k)
}
```
## Incident 2020
### All RRT
```{r, echo=FALSE, results='asis'}
#create tables
list.of.dts<-lapply(df[demogs], function(x) table(x, df$source_inc_rrt, useNA="ifany"))
#printing
for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=paste("Incident all RRT", demogs[i]), booktabs=TRUE) %>%
          kable_styling()
  print(k)
}
```
### Dialysis
```{r, echo=FALSE, results='asis'}
#create tables
list.of.dts<-lapply(df[demogs], function(x) table(x, df$source_inc_dialysis, useNA="ifany"))
#printing
for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=paste("Incident dialysis", demogs[i]), booktabs=TRUE) %>%
          kable_styling()
  print(k)
}
```
### Transplant
```{r, echo=FALSE, results='asis'}
#create tables
list.of.dts<-lapply(df[demogs], function(x) table(x, df$source_inc_transplant, useNA="ifany"))
#printing
for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=paste("Incident transplant", demogs[i]), booktabs=TRUE) %>%
          kable_styling()
  print(k)
}
```


# Figure 2 Numbers for Venn digram, sens and spec
## Prevalent at 1/1/2020
### All RRT
```{r, echo=FALSE, results='asis'}
table(df$source_prev_rrt, useNA="ifany") %>%
  kbl(col.names=c("Source", "N")) %>%
  kable_styling()
```
### Dialysis
```{r, echo=FALSE, results='asis'}
table(df$source_prev_dialysis, useNA="ifany") %>%
  kbl(col.names=c("Source", "N")) %>%
  kable_styling()
```
### Transplant
```{r, echo=FALSE, results='asis'}
table(df$source_prev_transplant, useNA="ifany") %>%
  kbl(col.names=c("Source", "N")) %>%
  kable_styling()
```
## Incident 2020
### All RRT
```{r, echo=FALSE, results='asis'}
table(df$source_inc_rrt, useNA="ifany") %>%
  kbl(col.names=c("Source", "N")) %>%
  kable_styling()
```
### Dialysis
```{r, echo=FALSE, results='asis'}
table(df$source_inc_dialysis, useNA="ifany") %>%
  kbl(col.names=c("Source", "N")) %>%
  kable_styling()
```
### Transplant
```{r, echo=FALSE, results='asis'}
table(df$source_inc_transplant, useNA="ifany") %>%
  kbl(col.names=c("Source", "N")) %>%
  kable_styling()
```

# Extra - how many are the wrong modality?
##prevalent 1/1/2020
### UKRR and PC
```{r, echo=FALSE, results='asis'}
table(df[df$source_prev_rrt %in% c("UP", "UPS"),]$ukrr_dtx,
      df[df$source_prev_rrt %in% c("UP", "UPS"),]$latest_rrt_status) %>%
  kbl() %>%
  add_header_above(c("UKRR", "Primary care" = 3)) %>%
  kable_styling()
```
### UKRR and SC
```{r, echo=FALSE, results='asis'}
k<-table(df[df$source_prev_rrt %in% c("US", "UPS"),]$ukrr_dtx,
      df[df$source_prev_rrt %in% c("US", "UPS"),]$incident_rrt_status_secondary) 
k2<-k%>%
  kbl() %>%
  add_header_above(c("UKRR", "Secondary care" = ncol(k))) %>%
  kable_styling()
```
## incident 2020
### UKRR and PC
```{r, echo=FALSE, results='asis'}
k<-table(df[df$source_inc_rrt %in% c("UP", "UPS"),]$ukrr_dtx,
      df[df$source_inc_rrt %in% c("UP", "UPS"),]$incident_rrt_status)  
k2<-k%>%
  kbl() %>%
  add_header_above(c("UKRR", "Primary care" = ncol(k))) %>%
  kable_styling()
```
### UKRR and SC
```{r, echo=FALSE, results='asis'}
k<-table(df[df$source_inc_rrt %in% c("US", "UPS"),]$ukrr_dtx,
      df[df$source_inc_rrt %in% c("US", "UPS"),]$incident_rrt_status_secondary)  
k2<-k%>%
  kbl() %>%
  add_header_above(c("UKRR", "Secondary care" = ncol(k))) %>%
  kable_styling()
```


# Figure 4 - CKD codes in PC/SC care for people in UKRR and not PC/SC, by modality
## Prevalent 1/1/2020
### in UKRR and not PC
```{r, echo=FALSE, results='asis'}
k<-table(df[df$source_prev_rrt %in% c("US", "U"),]$ukrr_dtx,
      df[df$source_prev_rrt %in% c("US", "U"),]$latest_renal_status)  
k2<-k%>%
  kbl() %>%
  add_header_above(c("UKRR", "Primary care" = ncol(k))) %>%
  kable_styling()
```
### in UKRR and not SC
```{r, echo=FALSE, results='asis'}
k<-table(df[df$source_prev_rrt %in% c("UP", "U"),]$ukrr_dtx,
      df[df$source_prev_rrt %in% c("UP", "U"),]$latest_ckd_status_secondary) 
k2<-k%>%
  kbl() %>%
  add_header_above(c("UKRR", "Secondary care" = ncol(k))) %>%
  kable_styling()
```
## Incident 2020
### in UKRR and not PC
```{r, echo=FALSE, results='asis'}
k<-table(df[df$source_inc_rrt %in% c("US", "U"),]$ukrr_dtx,
      df[df$source_inc_rrt %in% c("US", "U"),]$latest_renal_status)
k2<-k%>%
  kbl() %>%
  add_header_above(c("UKRR", "Primary care" = ncol(k))) %>%
  kable_styling()
```
### in UKRR and not SC
```{r, echo=FALSE, results='asis'}
k<-table(df[df$source_inc_rrt %in% c("UP", "U"),]$ukrr_dtx,
      df[df$source_inc_rrt %in% c("UP", "U"),]$latest_ckd_status_secondary) 
k2<-k %>%
  kbl() %>%
  add_header_above(c("UKRR", "Secondary care" = ncol(k))) %>%
  kable_styling()
print(k)
```

       

