---
title: "Data for tables and figures - paeds only"
author: "Shalini Santhakumaran"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

This document describes the cohorts of paediatric patients prevalent to renal replacement therapy (RRT) at 1st January 2020 using 3 sources of data: the UK Renal Registry (UKRR), primary care data from TPP practices (PC) and secondary care inpatient and outpatient data (SC).  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Load libraries
library(tidyverse)
library(knitr)
library(kableExtra)

# Create directory if it doesnt exist
fs::dir_create(here::here("output", "notebooks"))

demogs<-c("age_band2", "sex")

#read in dataset
#pick up 1st Jan cohort as latest renal status is on or before index date
 df<-read.csv(here::here("output", "input_paeds.csv.gz")) 
 

#add variable for dialysis/tx for UKRR
df<- df %>%
  mutate(ukrr_dtx = ifelse(ukrr_2019_mod %in% c("ICHD", "HD", "HHD", "PD") , "Dialysis", ifelse(ukrr_2019_mod == "Tx", "Transplant", "Not on RRT"))) %>%
  mutate(region = replace(region, region=="", "Missing")) 
# order factor
df$ukrr_dtx<-factor(df$ukrr_dtx, levels=c("Not on RRT", "Dialysis", "Transplant"))

# add variable for source
df <- df %>%
  mutate(source_prev_rrt = case_when(ukrr_2019==1 & 
                                      latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "UPS",
                                      ukrr_2019!=1 & 
                                      latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "PS",
                                     ukrr_2019==1 & 
                                      !(latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown")) &
                                      latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "US",
                                     ukrr_2019==1 & 
                                      latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      !(latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown")) ~ "UP",
                                     ukrr_2019==1 & 
                                      !(latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown")) &
                                      !(latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown")) ~ "U",
                                     ukrr_2019!=1 & 
                                      latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      !(latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown")) ~ "P",
                                     ukrr_2019!=1 & 
                                      !(latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown")) &
                                      latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "S")) %>%
  mutate(source_prev_dialysis = case_when(ukrr_dtx=="Dialysis" & 
                                      latest_rrt_status  =="Dialysis" &
                                      latest_rrt_status_secondary =="Dialysis" ~ "UPS",
                                      ukrr_dtx!="Dialysis" & 
                                      latest_rrt_status  =="Dialysis" &
                                      latest_rrt_status_secondary =="Dialysis" ~ "PS",
                                     ukrr_dtx=="Dialysis" & 
                                      !(latest_rrt_status  =="Dialysis") &
                                      latest_rrt_status_secondary =="Dialysis" ~ "US",
                                     ukrr_dtx=="Dialysis" & 
                                      latest_rrt_status  =="Dialysis" &
                                      !(latest_rrt_status_secondary =="Dialysis") ~ "UP",
                                     ukrr_dtx=="Dialysis" & 
                                      !(latest_rrt_status  =="Dialysis") &
                                      !(latest_rrt_status_secondary =="Dialysis") ~ "U",
                                     ukrr_dtx!="Dialysis" & 
                                      latest_rrt_status  =="Dialysis" &
                                      !(latest_rrt_status_secondary =="Dialysis") ~ "P",
                                     ukrr_dtx!="Dialysis" & 
                                      !(latest_rrt_status  =="Dialysis") &
                                      latest_rrt_status_secondary =="Dialysis" ~ "S")) %>%  
    mutate(source_prev_transplant = case_when(ukrr_dtx=="Transplant" & 
                                      latest_rrt_status  =="Transplant" &
                                      latest_rrt_status_secondary =="Transplant" ~ "UPS",
                                      ukrr_dtx!="Transplant" & 
                                      latest_rrt_status  =="Transplant" &
                                      latest_rrt_status_secondary =="Transplant" ~ "PS",
                                     ukrr_dtx=="Transplant" & 
                                      !(latest_rrt_status  =="Transplant") &
                                      latest_rrt_status_secondary =="Transplant" ~ "US",
                                     ukrr_dtx=="Transplant" & 
                                      latest_rrt_status  =="Transplant" &
                                      !(latest_rrt_status_secondary =="Transplant") ~ "UP",
                                     ukrr_dtx == "Transplant" & 
                                      !(latest_rrt_status  =="Transplant") &
                                      !(latest_rrt_status_secondary =="Transplant") ~ "U",
                                     ukrr_dtx!="Transplant" & 
                                      latest_rrt_status  =="Transplant" &
                                      !(latest_rrt_status_secondary =="Transplant") ~ "P",
                                     ukrr_dtx!="Transplant" & 
                                      !(latest_rrt_status  =="Transplant") &
                                      latest_rrt_status_secondary =="Transplant" ~ "S")) 
  
# order factor
source_levels<-c("UPS", "UP", "US", "PS", "U", "P", "S")
df$source_prev_rrt<-factor(df$source_prev_rrt, levels=source_levels)
df$source_prev_dialysis<-factor(df$source_prev_dialysis, levels=source_levels)
df$source_prev_transplant<-factor(df$source_prev_transplant, levels=source_levels)



#apply disclosure control - redact if less than 7, then round to nearest 5
disclosure <-function(x, whichmargins=0){
  redact <- 0<x & x<=7  #flag redacted values for later
  x[redact] <- 0  # set redacted to 0
  x<-round(x/5)*5 # round to nearest 5
  if(whichmargins!=0){
    x<-addmargins(x,whichmargins) #add totals
  }
  if(whichmargins==1){
    redact<-rbind(redact, rep(FALSE,ncol(redact))) #add extra row to redaction, otherwise won't blank correctly
  }
  if(whichmargins==2){
    redact<-cbind(redact, rep(FALSE,nrow(redact))) #add extra col to redaction, otherwise won't blank correctly
  }
  x[redact]<-"<=7"  # flag redacted values
  return(x)
}

```



Agreement is indicated by the data source, defined as UPS = in the **U**KRR, **P**rimary care and **S**econdary care data; UP = in **U**KRR and **P**rimary care only; U = in **U**KRR only etc.  

## Prevalent at 1/1/2020, overall and by modality
### All RRT
```{r, echo=FALSE, results='asis', error=TRUE}
disclosure(table(df$source_prev_rrt, useNA="ifany"),0) %>%
  kbl(col.names=c("Source", "N")) %>%
  kable_styling()
```
### Dialysis
```{r, echo=FALSE, results='asis', error=TRUE}
disclosure(table(df$source_prev_rrt, useNA="ifany"),0) %>%
  kbl(col.names=c("Source", "N")) %>%
  kable_styling()
```
### Transplant
```{r, echo=FALSE, results='asis', error=TRUE}
disclosure(table(df$source_prev_rrt, useNA="ifany"),0) %>%
  kbl(col.names=c("Source", "N")) %>%
  kable_styling()
```
