---
title: "Data for tables and figures"
author: "Shalini Santhakumaran"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

This document describes the cohorts of patients prevalent to renal replacement therapy (RRT) at 1st January 2020, and incident to renal replacement therapy during 2020 using 3 sources of data: the UK Renal Registry (UKRR), primary care data from TPP practices (PC) and secondary care inpatient and outpatient data (SC).  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Load libraries
library(tidyverse)
library(knitr)
library(kableExtra)

# Create directory if it doesnt exist
fs::dir_create(here::here("output", "notebooks"))

demogs<-c("age_band", "sex", "ethnicity", "imd", "diabetes", "hypertension",  "at_risk", "region")

#read in dataset
#pick up 1st Jan cohort as latest renal status is on or before index date
 df<-read.csv(here::here("output", "ukrr", "input_ukrr_2020-01-01.csv.gz")) 
 
 #only keep incident data for 2020 for PC and SC
df<-df %>%
  mutate(incident_rrt_status=ifelse(grepl("2020",incident_rrt_date), incident_rrt_status, "" )) %>%
  mutate(incident_rrt_status_secondary=ifelse(grepl("2020",incident_rrt_date_secondary), incident_rrt_status_secondary, "" ))

#add variable for dialysis/tx for UKRR
df<- df %>%
  mutate(ukrr_dtx = ifelse(ukrr_2019_mod %in% c("ICHD", "HD", "HHD", "PD") , "Dialysis", ifelse(ukrr_2019_mod == "Tx", "Transplant", "Not on RRT"))) %>%
  mutate(ukrr_dtx_inc = ifelse(ukrr_inc2020_mod %in% c("ICHD", "HD", "HHD", "PD") , "Dialysis", ifelse(ukrr_inc2020_mod == "Tx", "Transplant", "Not on RRT"))) %>%
  mutate(region = replace(region, region=="", "Missing"))
# order factor
df$ukrr_dtx<-factor(df$ukrr_dtx, levels=c("Not on RRT", "Dialysis", "Transplant"))
df$ukrr_dtx_inc<-factor(df$ukrr_dtx_inc, levels=c("Not on RRT", "Dialysis", "Transplant"))

# add variable for source
df <- df %>%
  mutate(source_prev_rrt = case_when(ukrr_2019==1 & 
                                      latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "UPS",
                                      ukrr_2019!=1 & 
                                      latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "PS",
                                     ukrr_2019==1 & 
                                      !(latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown")) &
                                      latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "US",
                                     ukrr_2019==1 & 
                                      latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      !(latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown")) ~ "UP",
                                     ukrr_2019==1 & 
                                      !(latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown")) &
                                      !(latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown")) ~ "U",
                                     ukrr_2019!=1 & 
                                      latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      !(latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown")) ~ "P",
                                     ukrr_2019!=1 & 
                                      !(latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown")) &
                                      latest_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "S")) %>%
  mutate(source_prev_dialysis = case_when(ukrr_dtx=="Dialysis" & 
                                      latest_rrt_status  =="Dialysis" &
                                      latest_rrt_status_secondary =="Dialysis" ~ "UPS",
                                      ukrr_dtx!="Dialysis" & 
                                      latest_rrt_status  =="Dialysis" &
                                      latest_rrt_status_secondary =="Dialysis" ~ "PS",
                                     ukrr_dtx=="Dialysis" & 
                                      !(latest_rrt_status  =="Dialysis") &
                                      latest_rrt_status_secondary =="Dialysis" ~ "US",
                                     ukrr_dtx=="Dialysis" & 
                                      latest_rrt_status  =="Dialysis" &
                                      !(latest_rrt_status_secondary =="Dialysis") ~ "UP",
                                     ukrr_dtx=="Dialysis" & 
                                      !(latest_rrt_status  =="Dialysis") &
                                      !(latest_rrt_status_secondary =="Dialysis") ~ "U",
                                     ukrr_dtx!="Dialysis" & 
                                      latest_rrt_status  =="Dialysis" &
                                      !(latest_rrt_status_secondary =="Dialysis") ~ "P",
                                     ukrr_dtx!="Dialysis" & 
                                      !(latest_rrt_status  =="Dialysis") &
                                      latest_rrt_status_secondary =="Dialysis" ~ "S")) %>%  
    mutate(source_prev_transplant = case_when(ukrr_dtx=="Transplant" & 
                                      latest_rrt_status  =="Transplant" &
                                      latest_rrt_status_secondary =="Transplant" ~ "UPS",
                                      ukrr_dtx!="Transplant" & 
                                      latest_rrt_status  =="Transplant" &
                                      latest_rrt_status_secondary =="Transplant" ~ "PS",
                                     ukrr_dtx=="Transplant" & 
                                      !(latest_rrt_status  =="Transplant") &
                                      latest_rrt_status_secondary =="Transplant" ~ "US",
                                     ukrr_dtx=="Transplant" & 
                                      latest_rrt_status  =="Transplant" &
                                      !(latest_rrt_status_secondary =="Transplant") ~ "UP",
                                     ukrr_dtx == "Transplant" & 
                                      !(latest_rrt_status  =="Transplant") &
                                      !(latest_rrt_status_secondary =="Transplant") ~ "U",
                                     ukrr_dtx!="Transplant" & 
                                      latest_rrt_status  =="Transplant" &
                                      !(latest_rrt_status_secondary =="Transplant") ~ "P",
                                     ukrr_dtx!="Transplant" & 
                                      !(latest_rrt_status  =="Transplant") &
                                      latest_rrt_status_secondary =="Transplant" ~ "S")) %>%  
    mutate(source_inc_rrt = case_when(ukrr_inc2020==1 & 
                                      incident_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      incident_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "UPS",
                                      ukrr_inc2020!=1 & 
                                      incident_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      incident_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "PS",
                                     ukrr_inc2020==1 & 
                                      !(incident_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown")) &
                                      incident_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "US",
                                     ukrr_inc2020==1 & 
                                      incident_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      !(incident_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown")) ~ "UP",
                                     ukrr_inc2020==1 & 
                                      !(incident_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown")) &
                                      !(incident_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown")) ~ "U",
                                     ukrr_inc2020!=1 & 
                                      incident_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown") &
                                      !(incident_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown")) ~ "P",
                                     ukrr_inc2020!=1 & 
                                      !(incident_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown")) &
                                      incident_rrt_status_secondary %in% c("Dialysis","Transplant", "RRT_unknown") ~ "S")) %>%
  mutate(source_inc_dialysis = case_when(ukrr_dtx_inc=="Dialysis" & 
                                      incident_rrt_status  =="Dialysis" &
                                      incident_rrt_status_secondary =="Dialysis" ~ "UPS",
                                      ukrr_dtx_inc!="Dialysis" & 
                                      incident_rrt_status  =="Dialysis" &
                                      incident_rrt_status_secondary =="Dialysis" ~ "PS",
                                     ukrr_dtx_inc=="Dialysis" & 
                                      !(incident_rrt_status  =="Dialysis") &
                                      incident_rrt_status_secondary =="Dialysis" ~ "US",
                                     ukrr_dtx_inc=="Dialysis" & 
                                      incident_rrt_status  =="Dialysis" &
                                      !(incident_rrt_status_secondary =="Dialysis") ~ "UP",
                                     ukrr_dtx_inc == "Dialysis" & 
                                      !(incident_rrt_status  =="Dialysis") &
                                      !(incident_rrt_status_secondary =="Dialysis") ~ "U",
                                     ukrr_dtx_inc!="Dialysis" & 
                                      incident_rrt_status  =="Dialysis" &
                                      !(incident_rrt_status_secondary =="Dialysis") ~ "P",
                                     ukrr_dtx_inc!="Dialysis" & 
                                      !(incident_rrt_status  =="Dialysis") &
                                      incident_rrt_status_secondary =="Dialysis" ~ "S")) %>%  
    mutate(source_inc_transplant = case_when(ukrr_dtx_inc=="Transplant" & 
                                      incident_rrt_status  =="Transplant" &
                                      incident_rrt_status_secondary =="Transplant" ~ "UPS",
                                      ukrr_dtx_inc!="Transplant" & 
                                      incident_rrt_status  =="Transplant" &
                                      incident_rrt_status_secondary =="Transplant" ~ "PS",
                                     ukrr_dtx_inc=="Transplant" & 
                                      !(incident_rrt_status  =="Transplant") &
                                      incident_rrt_status_secondary =="Transplant" ~ "US",
                                     ukrr_dtx_inc=="Transplant" & 
                                      incident_rrt_status  =="Transplant" &
                                      !(incident_rrt_status_secondary =="Transplant") ~ "UP",
                                     ukrr_dtx_inc == "Transplant" & 
                                      !(incident_rrt_status  =="Transplant") &
                                      !(incident_rrt_status_secondary =="Transplant") ~ "U",
                                     ukrr_dtx_inc!="Transplant" & 
                                      incident_rrt_status  =="Transplant" &
                                      !(incident_rrt_status_secondary =="Transplant") ~ "P",
                                     ukrr_dtx_inc!="Transplant" & 
                                      !(incident_rrt_status  =="Transplant") &
                                      incident_rrt_status_secondary =="Transplant" ~ "S"))

# pick out latest RRT code for primary care
df<-df %>%
    mutate(latest_rrt_code = ifelse(latest_rrt_status=="None", "None",
                                  ifelse(latest_rrt_status=="Dialysis", dialysis_code,
                                      ifelse(latest_rrt_status=="Transplant", kidney_tx_code,
                                             ifelse(latest_rrt_status=="RRT_unknown", RRT_code,"Uncategorised")))))


# order factor
source_levels<-c("UPS", "UP", "US", "PS", "U", "P", "S")
df$source_prev_rrt<-factor(df$source_prev_rrt, levels=source_levels)
df$source_prev_dialysis<-factor(df$source_prev_dialysis, levels=source_levels)
df$source_prev_transplant<-factor(df$source_prev_transplant, levels=source_levels)
df$source_inc_rrt<-factor(df$source_inc_rrt, levels=source_levels)
df$source_inc_dialysis<-factor(df$source_inc_dialysis, levels=source_levels)
df$source_inc_transplant<-factor(df$source_inc_transplant, levels=source_levels)

#read in codelists so we can tabulate the names
dialysis_codelist<-unique(read.csv(here::here("codelists", "opensafely-dialysis.csv"))[,1:2])
kidney_tx_codelist<-unique(read.csv(here::here("codelists", "opensafely-kidney-transplant.csv"))[,1:2])
all.codelists<-list(dialysis_codelist, kidney_tx_codelist )


# function to filter dataset to exclude patients with no RRT in either OS or UKRR as this is what we want to show pct by
keepRRT<-function(x){
  filter(x , ukrr_prev==1 | latest_renal_status %in% c("Dialysis", "Transplant", "RRT_unknown"))
}
keepRRT_nockd<-function(x){
  filter(x , ukrr_prev==1 | latest_rrt_status %in% c("Dialysis", "Transplant", "RRT_unknown"))
}

#apply disclosure control - redact if less than 7, then round to nearest 5
disclosure <-function(x, whichmargins=0){
  redact <- x<=7  #flag redacted values for later
  x[redact] <- 0  # set redacted to 0
  x<-round(x/5)*5 # round to nearest 5
  if(whichmargins!=0){
    x<-addmargins(x,whichmargins) #add totals
  }
  if(whichmargins==1){
    redact<-rbind(redact, rep(FALSE,ncol(redact))) #add extra row to redaction, otherwise won't blank correctly
  }
  if(whichmargins==2){
    redact<-cbind(redact, rep(FALSE,nrow(redact))) #add extra col to redaction, otherwise won't blank correctly
  }
  x[redact]<-"<=7"  # flag redacted values
  return(x)
}

```

# Section 1 Chacteristics of the study population
##  Prevalent at 1/1/2020, by data source and modality
### UKRR
```{r, echo=FALSE, results='asis', error=TRUE}
#create tables
list.of.dts<-lapply(df[df$ukrr_2019==1, demogs],
    function(x) disclosure(table(x, df[df$ukrr_2019==1,]$ukrr_dtx, useNA="ifany"),2))
#printing
for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=paste("Prevalent UKRR", demogs[i]), booktabs=TRUE) %>%
          kable_styling()
  print(k)
}
```
### Primary care
```{r, echo=FALSE, results='asis', error=TRUE}
#create tables
list.of.dts<-lapply(df[df$latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown"), demogs],
       function(x) disclosure(table(x, df[df$latest_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown"),]$latest_rrt_status, useNA="ifany"),2))
#printing
for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=paste("Prevalent primary care", demogs[i]), booktabs=TRUE) %>%
          kable_styling()
  print(k)
}
```
### Secondary care
```{r, echo=FALSE, results='asis', error=TRUE}
#create tables
list.of.dts<-lapply(df[df$latest_rrt_status_secondary  %in%  c("Dialysis", "Transplant", "RRT_unknown"), demogs],
       function(x) 
         disclosure(
           table(x, df[df$latest_rrt_status_secondary  %in%  c("Dialysis", "Transplant", "RRT_unknown"),]$latest_rrt_status_secondary, useNA="ifany")
           ,2))
#printing
for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=paste("Prevalent secondary care", demogs[i]), booktabs=TRUE) %>%
          kable_styling()
  print(k)
}
```

## Incident in 2020, by data source and modality
### UKRR
```{r, echo=FALSE, results='asis', error=TRUE}
#create tables
list.of.dts<-lapply(df[df$ukrr_inc2020==1, demogs],
       function(x) disclosure(table(x, df[df$ukrr_inc2020==1,]$ukrr_dtx_inc, useNA="ifany"),2))
#printing
for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=paste("Incident UKRR", demogs[i]), booktabs=TRUE) %>%
          kable_styling()
  print(k)
}
```
### Primary care 
```{r, echo=FALSE, results='asis', error=TRUE}
#create tables
list.of.dts<-lapply(df[df$incident_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown"), demogs],
       function(x) disclosure(table(x, df[df$incident_rrt_status  %in%  c("Dialysis", "Transplant", "RRT_unknown"),]$incident_rrt_status, useNA="ifany"),2))
#printing
for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=paste("Incident primary care", demogs[i]), booktabs=TRUE) %>%
          kable_styling()
  print(k)
}
```

### Secondary care
```{r, echo=FALSE, results='asis', error=TRUE}
#create tables
list.of.dts<-lapply(df[df$incident_rrt_status_secondary  %in%  c("Dialysis", "Transplant", "RRT_unknown"), demogs],
       function(x) disclosure(
         table(x, df[df$incident_rrt_status_secondary  %in%  c("Dialysis", "Transplant", "RRT_unknown"),]$incident_rrt_status_secondary, useNA="ifany")
         ,2))
#printing
for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=paste("Incident secondary care", demogs[i]), booktabs=TRUE) %>%
          kable_styling()
  print(k)
}
```

### add median age
```{r, echo=FALSE, results='asis', error=TRUE}
#create tables

med.age<-function(x){
  x <- enquo(x)
  df%>%
    group_by(!!x)%>% 
    summarise(age = quantile(age, c(0.25, 0.5, 0.75), na.rm=TRUE), q = c(0.25, 0.5, 0.75)) %>%
    spread(q, age)
  }

  k<- med.age(ukrr_dtx) %>%
          kbl(caption="Prevalent UKRR", booktabs=TRUE) %>%
          kable_styling()
  print(k)

    k<- med.age(latest_rrt_status ) %>%
          kbl(caption="Prevalent primary care", booktabs=TRUE) %>%
          kable_styling()
  print(k)
  
    k<- med.age(latest_rrt_status_secondary) %>%
          kbl(caption="Prevalent secondary care", booktabs=TRUE) %>%
          kable_styling()
  print(k)

```


# Section 2 Overall agreement  

Agreement is indicated by the data source, defined as UPS = in the **U**KRR, **P**rimary care and **S**econdary care data; UP = in **U**KRR and **P**rimary care only; U = in **U**KRR only etc.  

## Prevalent at 1/1/2020, overall and by modality
### All RRT
```{r, echo=FALSE, results='asis', error=TRUE}
disclosure(table(df$source_prev_rrt, useNA="ifany"),0) %>%
  kbl(col.names=c("Source", "N")) %>%
  kable_styling()
```
### Dialysis
```{r, echo=FALSE, results='asis', error=TRUE}
disclosure(table(df$source_prev_dialysis, useNA="ifany"),0) %>%
  kbl(col.names=c("Source", "N")) %>%
  kable_styling()
```
### Transplant
```{r, echo=FALSE, results='asis', error=TRUE}
disclosure(table(df$source_prev_transplant, useNA="ifany"),0) %>%
  kbl(col.names=c("Source", "N")) %>%
  kable_styling()
```
## Incident 2020, overall and by modality
### All RRT
```{r, echo=FALSE, results='asis', error=TRUE}
disclosure(table(df$source_inc_rrt, useNA="ifany"),0) %>%
  kbl(col.names=c("Source", "N")) %>%
  kable_styling()
```
### Dialysis
```{r, echo=FALSE, results='asis', error=TRUE}
disclosure(table(df$source_inc_dialysis, useNA="ifany"),0) %>%
  kbl(col.names=c("Source", "N")) %>%
  kable_styling()
```
### Transplant
```{r, echo=FALSE, results='asis', error=TRUE}
disclosure(table(df$source_inc_transplant, useNA="ifany"),0) %>%
  kbl(col.names=c("Source", "N")) %>%
  kable_styling()
```

# Section 3 Agreement by demographic characteristics
## Prevalent at 1/1/2020, overall and by modality
### All RRT

```{r, echo=FALSE, results='asis', error=TRUE}
#create tables
list.of.dts<-lapply(df[demogs], function(x) disclosure(table(x, df$source_prev_rrt, useNA="ifany"),0))
#printing
for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=paste("Prevalent, all RRT", demogs[i]), booktabs=TRUE) %>%
          kable_styling()
  print(k)
}
```
### Dialysis
```{r, echo=FALSE, results='asis', error=TRUE}
#create tables
list.of.dts<-lapply(df[demogs], function(x) disclosure(table(x, df$source_prev_dialysis, useNA="ifany"),0))
#printing
for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=paste("Prevalent dialysis", demogs[i]), booktabs=TRUE) %>%
          kable_styling()
  print(k)
}
```
### Transplant
```{r, echo=FALSE, results='asis', error=TRUE}
#create tables
list.of.dts<-lapply(df[demogs], function(x) disclosure(table(x, df$source_prev_transplant, useNA="ifany"),0))
#printing
for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=paste("Prevalent transplant", demogs[i]), booktabs=TRUE) %>%
          kable_styling()
  print(k)
}
```
## Incident 2020, overall and by modality
### All RRT
```{r, echo=FALSE, results='asis', error=TRUE}
#create tables
list.of.dts<-lapply(df[demogs], function(x) disclosure(table(x, df$source_inc_rrt, useNA="ifany"),0))
#printing
for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=paste("Incident all RRT", demogs[i]), booktabs=TRUE) %>%
          kable_styling()
  print(k)
}
```
### Dialysis
```{r, echo=FALSE, results='asis', error=TRUE}
#create tables
list.of.dts<-lapply(df[demogs], function(x) disclosure(table(x, df$source_inc_dialysis, useNA="ifany"),0))
#printing
for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=paste("Incident dialysis", demogs[i]), booktabs=TRUE) %>%
          kable_styling()
  print(k)
}
```
### Transplant
```{r, echo=FALSE, results='asis', error=TRUE}
#create tables
list.of.dts<-lapply(df[demogs], function(x) disclosure(table(x, df$source_inc_transplant, useNA="ifany"),0))
#printing
for(i in 1:length(list.of.dts)) {
  k<- list.of.dts[[i]] %>%
          kbl(caption=paste("Incident transplant", demogs[i]), booktabs=TRUE) %>%
          kable_styling()
  print(k)
}
```

# Section 4  - Agreement of modality  

Of people who are captured in UKRR and PC/SC, do the modalities agree?   

## Prevalent 1/1/2020
### UKRR and PC
```{r, echo=FALSE, results='asis', error=TRUE}
k<-disclosure(table(df[df$source_prev_rrt %in% c("UP", "UPS"),]$ukrr_dtx,
      df[df$source_prev_rrt %in% c("UP", "UPS"),]$latest_rrt_status),0)
k2<-k%>%
  kbl() %>%
  add_header_above(c("UKRR", "Primary care" = 3)) %>%
  kable_styling()
print(k2)
```
### UKRR and SC
```{r, echo=FALSE, results='asis', error=TRUE}
k<-disclosure(table(df[df$source_prev_rrt %in% c("US", "UPS"),]$ukrr_dtx,
      df[df$source_prev_rrt %in% c("US", "UPS"),]$incident_rrt_status_secondary),0) 
k2<-k%>%
  kbl() %>%
  add_header_above(c("UKRR", "Secondary care" = ncol(k))) %>%
  kable_styling()
print(k2)
```
## incident 2020
### UKRR and PC
```{r, echo=FALSE, results='asis', error=TRUE}
k<-disclosure(table(df[df$source_inc_rrt %in% c("UP", "UPS"),]$ukrr_dtx_inc,
      df[df$source_inc_rrt %in% c("UP", "UPS"),]$incident_rrt_status),0)  
k2<-k%>%
  kbl() %>%
  add_header_above(c("UKRR", "Primary care" = ncol(k))) %>%
  kable_styling()
print(k2)
```
### UKRR and SC
```{r, echo=FALSE, results='asis', error=TRUE}
k<-disclosure(table(df[df$source_inc_rrt %in% c("US", "UPS"),]$ukrr_dtx_inc,
      df[df$source_inc_rrt %in% c("US", "UPS"),]$incident_rrt_status_secondary),0)  
k2<-k%>%
  kbl() %>%
  add_header_above(c("UKRR", "Secondary care" = ncol(k))) %>%
  kable_styling()
print(k2)
```


# Section 5 - CKD codes in PC/SC care for people in UKRR and not PC/SC, by modality  
## Prevalent 1/1/2020
### in UKRR and not PC
```{r, echo=FALSE, results='asis', error=TRUE}
k<-disclosure(table(df[df$source_prev_rrt %in% c("US", "U"),]$ukrr_dtx,
      df[df$source_prev_rrt %in% c("US", "U"),]$latest_renal_status, exclude="Not on RRT")  ,0)
k2<-k%>%
  kbl() %>%
  add_header_above(c("UKRR", "Primary care" = ncol(k))) %>%
  kable_styling()
print(k2)
```
### in UKRR and not SC
```{r, echo=FALSE, results='asis', error=TRUE}
k<-disclosure(table(df[df$source_prev_rrt %in% c("UP", "U"),]$ukrr_dtx,
      df[df$source_prev_rrt %in% c("UP", "U"),]$latest_ckd_status_secondary) ,0)
k2<-k%>%
  kbl() %>%
  add_header_above(c("UKRR", "Secondary care" = ncol(k))) %>%
  kable_styling()
print(k2)
```
## Incident 2020
These might not make much sense as they could have been incident before or after 2020 in PC/SC.
### in UKRR and not PC
```{r, echo=FALSE, results='asis', error=TRUE}
k<-disclosure(table(df[df$source_inc_rrt %in% c("US", "U"),]$ukrr_dtx_inc,
      df[df$source_inc_rrt %in% c("US", "U"),]$latest_renal_status),0)
k2<-k%>%
  kbl() %>%
  add_header_above(c("UKRR", "Primary care" = ncol(k))) %>%
  kable_styling()
print(k2)
```
### in UKRR and not SC
```{r, echo=FALSE, results='asis', error=TRUE}
k<-disclosure(table(df[df$source_inc_rrt %in% c("UP", "U"),]$ukrr_dtx_inc,
      df[df$source_inc_rrt %in% c("UP", "U"),]$latest_ckd_status_secondary),0) 
k2<-k %>%
  kbl() %>%
  add_header_above(c("UKRR", "Secondary care" = ncol(k))) %>%
  kable_styling()
print(k2)
```

  # Section 6 Codes for people on RRT according to primary care data but not the UKRR
Below we look at what the most recent primary care codes are for these patients, by modality
```{r, echo=FALSE, results='asis'}

# Looking at median time since code
codes.table.3<-function(df,status, codelist){
  df %>%
    filter(source_prev_rrt %in% c("P", "PS") & latest_rrt_status==status) %>%
    group_by(latest_rrt_code)%>%
    summarise(n=n()) %>%
    left_join(codelist, by=setNames(names(codelist)[1], "latest_rrt_code")) %>%
    mutate(latest_rrt_code= replace(latest_rrt_code, latest_rrt_code %in% c("", NA) , "Missing")) %>%
      mutate(n=ifelse(n<=7,0,n)) %>% #first replace n<=7 with 0
      mutate(n=round(n/5)*5) %>%  #round to nearest 5
      mutate(Percentage = ifelse(n==0, NA, round(100*n/sum(n), 0))) %>%   #then calculate %s, blanking if n is 0
        arrange(desc(n))  %>% #descending sort by n
      mutate(n=ifelse(n<=5,"<=5",n)) #now replacing <=5 (which is now 0 anyway) with "<=5"
}

list.of.dts<- mapply(function(y, z) 
    codes.table.3(df, y, z), 
    c("Dialysis", "Transplant"),  all.codelists[1:2], SIMPLIFY=FALSE)

  for(j in 1:2){
   k<- list.of.dts[j]%>%
      

